---
title: "Analysis of neuropsych data"
output: html_notebook
---

Now we do some classification using rsFMRI. First, some data cleaning:

```{r}
# for replication purposes
set.seed(107)
source('~/ncr_notebooks/baseline_prediction//src/aux_functions.R')

# Philip pre-selected which structural scans to use
gf_fname = '~/data/baseline_prediction/gf_updated_02152017.csv'
gf = read.csv(gf_fname)
group_var = 'DX'
idx = gf$BASELINE == 'BASELINE'
gf = gf[idx,]
eval(parse(text=sprintf('groups = gf$\"%s\"', group_var)))
idx = groups=='NV' | groups=='ADHD'
groups = factor(groups[idx])
gf = gf[idx,]
groups_orig = groups
```

Now it's a matter of selecting the variables we want to test. As usual, we start with everything:

```{r}
phen_vars = c(#IQ
              'VIQ', 'PIQ', 'FSIQ',
              # Processing speed
              'PS_all',
              # CPT
              'N_of_omissions', 'PC_omissions', 'N_commissions', 'PC_commissions',
              'hit_RT', 'hit_RT_SE', 'variability_of_SE', 'D.prime', 'beta', 'N_perservations',
              'PC_perservations', 'hit_RT_block_change', 'hit_RT_SE_block_change',
              'hit_RT_ISI_change', 'hit_RT_SE_ISI_change',
              'omissions_t.scores_for_general_population',
              'commissions_t.scores_for_general_population', 
              'hit_RT_t.scores_for_general_population',
              'hit_RT_SE_t.scores_for_general_population',
              'variability_of_SE_t.scores_for_general_population',
              'D.prime_t.scores_for_general_population',
              'beta_t.scores_for_general_population',
              'perserverations_t.scores_for_general_population',
              'hit_RT_block_change_t.scores_for_general_population',
              'hit_RT_SE_block_change_t.scores_for_general_population',
              'hit_RT_ISI_change_t.scores_for_general_population',
              'X_hit_RT_SE_ISI_change_t.scores_for_general_population',
              'ADHD.confidence', 'focus', 'HI', 'sustained', 'vigilance',
              # WISC
              'DSF_raw', 'DSB_raw', 'DS_raw.1', 'SSF_raw', 'SSB_raw', 'SS_raw',
              'DSF_SS', 'DSB_SS', 'DS_SS', 'SSF_SS', 'SSD_SS', 'SS_SS',
              # SES
              'SES',
              # GAS
              'GAS',
              # FES
              'FES_Cohesion_T', 'FES_Expressiveness_T', 'FES_Conflict_T',
              'FES_Independence_T', 'FES_Achiev T', 'FES_Intellectual-Cultural_T',
              'FES_Active-Recreational_T', 'FES_Moral-Religious_T',
              'FES_Organization_T', 'FES_Control_T')
keep_me = c()
for (v in phen_vars) {
  keep_me = c(keep_me, which(colnames(gf) == v))
}
ldata = gf[, keep_me]
```

There are too many variables with NAs, and right now I don't have an intuition of whether imputation or removing subjects would work best. Let's do this within domain first, and evaluate in a per-domain basis.

# IQ

```{r}
conn_na = colSums(is.na(ldata)) / nrow(ldata)
phen_vars = c(#IQ
              'VIQ', 'PIQ', 'FSIQ')
keep_me = c()
for (v in phen_vars) {
  keep_me = c(keep_me, which(colnames(gf) == v))
}
ldata_orig = gf[, keep_me]
groups_orig = groups
```

First we try some imputation:
```{r}
ldata = ldata_orig
groups = groups_orig
preProcValues <- preProcess(ldata, method = c("medianImpute"))
ldata <- predict(preProcValues, ldata)
ncv = 5
nrepeatcv = 2
nsplits = 5
train_test_ratio = .85
root_fname = '~/tmp/iq_medianImputed_NVvsADHD'
# saving output
library(caret)
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
ncpus <- detectBatchCPUs()
njobs = ncpus - 1

run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')

inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
metric = "ROC"
default_preproc = c("center", 'scale')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

Just because some people might complain about a whole data imputation, let's try it within the cross-validation as well:

```{r}
ldata = ldata_orig
groups = groups_orig
root_fname = '~/tmp/iq_medianImputedCV_NVvsADHD'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale', 'medianImpute')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

Finally, let's just remove subjects with NAs:

```{r}
ldata = ldata_orig
groups = groups_orig
keep_me = rowSums(is.na(ldata)) == 0
ldata = ldata[keep_me, ]
groups = groups[keep_me]
root_fname = '~/tmp/iq_NAremoved_NVvsADHD'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')

```
## PCA
Just because we'll need to PCA some of these data domains, let's construct the code base now:

```{r}
ldata = ldata_orig
groups = groups_orig

preProcValues <- preProcess(ldata, method = c("medianImpute"))
ldata <- predict(preProcValues, ldata)
root_fname = '~/tmp/iq_medianImputed_NVvsADHD_pca'
# saving output
library(caret)
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')

inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
metric = "ROC"
default_preproc = c('pca')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

Median impute during CV:
```{r}
ldata = ldata_orig
groups = groups_orig
root_fname = '~/tmp/iq_medianImputedCV_NVvsADHD_pca'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale', 'medianImpute', 'pca')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

And finally, pca after removing subjects:

```{r}
ldata = ldata_orig
groups = groups_orig
keep_me = rowSums(is.na(ldata)) == 0
ldata = ldata[keep_me, ]
groups = groups[keep_me]
root_fname = '~/tmp/iq_NAremoved_NVvsADHD_pca'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale', 'pca')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')

```

# PS, GAS, SES
I know these are not related, but they're all single variables, so at least this way we can do a multi-variate shenanigans.

```{r}
conn_na = colSums(is.na(ldata)) / nrow(ldata)
phen_vars = c('PS_all', 'SES', 'GAS')
keep_me = c()
for (v in phen_vars) {
  keep_me = c(keep_me, which(colnames(gf) == v))
}
ldata_orig = gf[, keep_me]
```

```{r}
ldata = ldata_orig
groups = groups_orig
preProcValues <- preProcess(ldata, method = c("medianImpute"))
ldata <- predict(preProcValues, ldata)
root_fname = '~/tmp/PSGASSES_medianImputed_NVvsADHD'
# saving output
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')

inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
metric = "ROC"
default_preproc = c("center", 'scale')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

Imputation within CV:

```{r}
ldata = ldata_orig
groups = groups_orig
root_fname = '~/tmp/PSGASSES_medianImputedCV_NVvsADHD'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale', 'medianImpute')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

Removing subjects with NA:

```{r}
ldata = ldata_orig
groups = groups_orig
keep_me = rowSums(is.na(ldata)) == 0
ldata = ldata[keep_me, ]
groups = groups[keep_me]
root_fname = '~/tmp/PSGASSES_NAremoved_NVvsADHD'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

## PCA
And just like before, we do the PCAs:

```{r}
ldata = ldata_orig
groups = groups_orig
preProcValues <- preProcess(ldata, method = c("medianImpute"))
ldata <- predict(preProcValues, ldata)
root_fname = '~/tmp/PSGASSES_medianImputed_NVvsADHD_pca'
# saving output
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')

inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
metric = "ROC"
default_preproc = c("center", 'scale', 'pca')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

```{r}
ldata = ldata_orig
groups = groups_orig
root_fname = '~/tmp/PSGASSES_medianImputedCV_NVvsADHD_pca'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale', 'medianImpute', 'pca')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

```{r}
ldata = ldata_orig
groups = groups_orig
keep_me = rowSums(is.na(ldata)) == 0
ldata = ldata[keep_me, ]
groups = groups[keep_me]
root_fname = '~/tmp/PSGASSES_NAremoved_NVvsADHD_pca'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale', 'pca')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

# CPT

```{r}
conn_na = colSums(is.na(ldata)) / nrow(ldata)
phen_vars = c('N_of_omissions', 'PC_omissions', 'N_commissions', 'PC_commissions',
              'hit_RT', 'hit_RT_SE', 'variability_of_SE', 'D.prime', 'beta', 'N_perservations',
              'PC_perservations', 'hit_RT_block_change', 'hit_RT_SE_block_change',
              'hit_RT_ISI_change', 'hit_RT_SE_ISI_change',
              'omissions_t.scores_for_general_population',
              'commissions_t.scores_for_general_population', 
              'hit_RT_t.scores_for_general_population',
              'hit_RT_SE_t.scores_for_general_population',
              'variability_of_SE_t.scores_for_general_population',
              'D.prime_t.scores_for_general_population',
              'beta_t.scores_for_general_population',
              'perserverations_t.scores_for_general_population',
              'hit_RT_block_change_t.scores_for_general_population',
              'hit_RT_SE_block_change_t.scores_for_general_population',
              'hit_RT_ISI_change_t.scores_for_general_population',
              'X_hit_RT_SE_ISI_change_t.scores_for_general_population',
              'ADHD.confidence', 'focus', 'HI', 'sustained', 'vigilance')
keep_me = c()
for (v in phen_vars) {
  keep_me = c(keep_me, which(colnames(gf) == v))
}
ldata_orig = gf[, keep_me]
```

```{r}
ldata = ldata_orig
groups = groups_orig
preProcValues <- preProcess(ldata, method = c("medianImpute"))
ldata <- predict(preProcValues, ldata)
root_fname = '~/tmp/CPT_medianImputed_NVvsADHD'
# saving output
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')

inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
metric = "ROC"
default_preproc = c("center", 'scale')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

```{r}
ldata = ldata_orig
groups = groups_orig
root_fname = '~/tmp/CPT_medianImputedCV_NVvsADHD'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale', 'medianImpute')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

```{r}
ldata = ldata_orig
groups = groups_orig
keep_me = rowSums(is.na(ldata)) == 0
ldata = ldata[keep_me, ]
groups = groups[keep_me]
root_fname = '~/tmp/CPT_NAremoved_NVvsADHD'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

## PCA
And just like before, we do the PCAs:

```{r}
ldata = ldata_orig
groups = groups_orig
preProcValues <- preProcess(ldata, method = c("medianImpute"))
ldata <- predict(preProcValues, ldata)
root_fname = '~/tmp/CPT_medianImputed_NVvsADHD_pca'
# saving output
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')

inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
metric = "ROC"
default_preproc = c("center", 'scale', 'pca')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

```{r}
ldata = ldata_orig
groups = groups_orig
root_fname = '~/tmp/CPT_medianImputedCV_NVvsADHD_pca'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale', 'medianImpute', 'pca')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```

```{r}
ldata = ldata_orig
groups = groups_orig
keep_me = rowSums(is.na(ldata)) == 0
ldata = ldata[keep_me, ]
groups = groups[keep_me]
root_fname = '~/tmp/CPT_NAremoved_NVvsADHD_pca'
sink(sprintf('%s.log', root_fname), append=FALSE, split=TRUE)
run_models = c('rndForest', 'lr', 'rsvm', 'xgb') # c('rndForest', 'lr', 'lsvm', 'rsvm', 'xgb', 'gbm')
inTrain = createDataPartition(y = groups,
                              times= nsplits,
                              p=train_test_ratio)
default_preproc = c("center", 'scale', 'pca')
default_options = c()
ctrl_cv <- trainControl(method = "repeatedcv",
                        number = ncv,
                        repeats = nrepeatcv,
                        classProbs = TRUE,
                        returnData = FALSE,
                        summaryFunction = twoClassSummary,
                        preProcOptions = default_options,
                        search='grid')

source('~/ncr_notebooks/baseline_prediction/src/do_classification.R')
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
sink()
```



