---
title: "Data descriptives"
output: html_notebook
---

Philip asked for the usual descriptives for each dataset. Let's do it for the usual targets, and then export them all to a nice table.

# DTI

```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
rm_me = (tract_data$fa_avg < .4 | tract_data$ad_avg < 1.18 | tract_data$rd_avg > .65 | tract_data$rd_avg < .5 |
         tract_data$norm.trans > .45 | tract_data$norm.rot > .008 | tract_data$goodSlices < 45 | tract_data$goodSlices > 70)
print(sprintf('Reducing from %d to %d scans', nrow(tract_data), nrow(tract_data)-sum(rm_me)))
tract_data = tract_data[!rm_me, ]
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf$MRN, tract_data$MRN)
merged = mergeOnClosestDate(gf, tract_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
print(sprintf('Reducing from %d to %d scans', nrow(merged), nrow(merged)-sum(rm_me)))
merged = merged[!rm_me, ]
phen_vars = c(which(grepl("^FA_", colnames(merged))),
              which(grepl("^AD_", colnames(merged))),
              which(grepl("^RD_", colnames(merged))),
              which(grepl("^MO_", colnames(merged)))
              )
X = merged[, phen_vars]
```


```{r}
nvVSadhd_test = function(X, merged, dtype) {
  y = as.character(merged$DX_BASELINE)
  y[y != 'NV'] = 'ADHD'
  y = factor(y, levels=c('NV', 'ADHD'))
  modality <- apply(X, 2, function(dat) {dtype})
  
  # tstat <- apply(X, 2, function(dat) {
  #   +     summary(lm(dat ~ y + merged$age + merged$SEX))$coefficients[2, 3]})
  # pvals <- apply(X, 2, function(dat) {
  #   +     summary(lm(dat ~ y + merged$age + merged$SEX))$coefficients[2, 4]})
  
  # tstat <- apply(X, 2, function(dat) {
  #   +     summary(lm(dat ~ y))$coefficients[2, 3]})
  # pvals <- apply(X, 2, function(dat) {
  #   +     summary(lm(dat ~ y))$coefficients[2, 4]})
  
  tstat <- apply(X, 2, function(dat) {
    +     summary(lm(dat ~ y + merged$SEX))$coefficients[2, 3]})
  pvals <- apply(X, 2, function(dat) {
    +     summary(lm(dat ~ y + merged$SEX))$coefficients[2, 4]})
  
  res = cbind(modality, tstat, pvals)
  colnames(res) = c('dtype', 'nvVSadhd_tstat', 'nvVSadhd_pval')
  return(res)
}

nvVSadhd_lowVSadhd_high_test = function(X, merged) {
  y = as.character(merged$DX_BASELINE)
  y[y != 'NV'] = 'ADHD'
  # myc = kmeans(merged[y != 'NV',]$GAS, 2)
  # thresh = as.numeric(myc$centers[1] + diff(myc$centers)/2)
  # print(sprintf('Cutoff: %.2f', thresh))
  thresh = 68
  y[y == 'ADHD' & merged$GAS > thresh] = 'ADHD_high'
  y[y == 'ADHD'] = 'ADHD_low'
  y = factor(y, levels=c('NV', 'ADHD_low', 'ADHD_high'))
  
  # fstat <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~  + merged$age + merged$SEX + y)))[[1]][4][[3, 1]]})
  # pvals <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ merged$age + merged$SEX + y)))[[1]][5][[3, 1]]})
  
  # fstat <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ y)))[[1]][4][[1, 1]]})
  # pvals <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ y)))[[1]][5][[1, 1]]})
  
  fstat <- apply(X, 2, function(dat) {
    +     summary(aov(lm(dat ~ merged$SEX + y)))[[1]][4][[2, 1]]})
  pvals <- apply(X, 2, function(dat) {
    +     summary(aov(lm(dat ~ merged$SEX + y)))[[1]][5][[2, 1]]})
  
  res = cbind(fstat, pvals)
  colnames(res) = c('nvVSadhdlowVSadhdhigh_Fstat', 'nvVSadhdlowVSadhdhigh_pval')
  return(res)
}

outcome_test = function(X, merged) {
  y = merged$inatt3_named
  # fstat <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ merged$age + merged$SEX + y)))[[1]][4][[3, 1]]})
  # pvals <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ merged$age + merged$SEX + y)))[[1]][5][[3, 1]]})

  # fstat <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ y)))[[1]][4][[1, 1]]})
  # pvals <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ y)))[[1]][5][[1, 1]]})
  
  fstat <- apply(X, 2, function(dat) {
    +     summary(aov(lm(dat ~ merged$SEX + y)))[[1]][4][[2, 1]]})
  pvals <- apply(X, 2, function(dat) {
    +     summary(aov(lm(dat ~ merged$SEX + y)))[[1]][5][[2, 1]]})
  
  res_inatt = cbind(fstat, pvals)
  colnames(res_inatt) = c('outcomeInatt3Grp_Fstat', 'outcomeInatt3Grp_pval')
  
  y = merged$HI3_named
  # fstat <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ merged$age + merged$SEX + y)))[[1]][4][[3, 1]]})
  # pvals <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ merged$age + merged$SEX + y)))[[1]][5][[3, 1]]})
  
  # fstat <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ y)))[[1]][4][[1, 1]]})
  # pvals <- apply(X, 2, function(dat) {
  #   +     summary(aov(lm(dat ~ y)))[[1]][5][[1, 1]]})
  
  fstat <- apply(X, 2, function(dat) {
    +     summary(aov(lm(dat ~ merged$SEX + y)))[[1]][4][[2, 1]]})
  pvals <- apply(X, 2, function(dat) {
    +     summary(aov(lm(dat ~ merged$SEX + y)))[[1]][5][[2, 1]]})
  
  res_hi = cbind(fstat, pvals)
  colnames(res_hi) = c('outcomeHI3Grp_Fstat', 'outcomeHI3Grp_pval')
  return(cbind(res_inatt, res_hi))
}

symptom_test = function(X, merged, thresh) {
  keep_me = merged$SX_inatt >= thresh & merged$SX_HI >= thresh
  X = X[keep_me, ]
  
  y = merged[keep_me,]$SX_inatt
  tstat <- apply(X, 2, function(dat) {
    +      summary(lm(dat ~ y + merged[keep_me,]$age + merged[keep_me,]$SEX))$coefficients[2, 3]})
  pvals <- apply(X, 2, function(dat) {
    +      summary(lm(dat ~ y + merged[keep_me,]$age + merged[keep_me,]$SEX))$coefficients[2, 4]})
  res_inatt = cbind(tstat, pvals)
  colnames(res_inatt) = c(sprintf('inattBE%d_tstat', thresh),
                          sprintf('inattBE%d_pval', thresh))
   
  y = merged[keep_me,]$SX_HI
  tstat <- apply(X, 2, function(dat) {
    +      summary(lm(dat ~ y + merged[keep_me,]$age + merged[keep_me,]$SEX))$coefficients[2, 3]})
  pvals <- apply(X, 2, function(dat) {
    +      summary(lm(dat ~ y + merged[keep_me,]$age + merged[keep_me,]$SEX))$coefficients[2, 4]})
  res_hi = cbind(tstat, pvals)
  colnames(res_hi) = c(sprintf('hiBE%d_tstat', thresh),
                          sprintf('hiBE%d_pval', thresh))
  
  return(cbind(res_inatt, res_hi))
}
```

Now that all functions are written, it's just a matter of combining their results:

```{r}
dti_descriptive = cbind(nvVSadhd_test(X, merged, 'DTI'),
                        nvVSadhd_lowVSadhd_high_test(X, merged),
                        outcome_test(X, merged),
                        symptom_test(X, merged, 0),
                        symptom_test(X, merged, 1),
                        symptom_test(X, merged, 2))
write.csv(dti_descriptive, file='~/tmp/oi.csv')
```

#PRS
```{r}
prs_data = read.csv('~/data/prs/PRS2017_noInversion_all.csv')
merged = merge(gf, prs_data, by='MRN')
phen_vars = which(grepl("^PROFILES", colnames(merged)))
keep_me = c()
for (v in phen_vars) {
  keep_me = c(keep_me, which(colnames(merged) == v))
}
X = merged[, keep_me]
prs_descriptive = cbind(nvVSadhd_test(X, merged, 'PRS'),
                        nvVSadhd_lowVSadhd_high_test(X, merged),
                        outcome_test(X, merged),
                        symptom_test(X, merged, 0),
                        symptom_test(X, merged, 1),
                        symptom_test(X, merged, 2))
```

# Demographics
This is still part of the clinical spreadsheet, so let's just do it without any extra import:

```{r}
# GAS doesn't make much sense, but I needed some other numerical variable to go with age so the functions would work
phen_vars = c(which(grepl("^age", colnames(gf))),
              which(grepl("^GAS", colnames(gf))))
X = gf[, phen_vars]
demo_descriptive = cbind(nvVSadhd_test(X, gf, 'demo'),
                        nvVSadhd_lowVSadhd_high_test(X, gf),
                        outcome_test(X, gf),
                        symptom_test(X, gf, 0),
                        symptom_test(X, gf, 1),
                        symptom_test(X, gf, 2))
```

# Geospatial
```{r}
geo_data = read.csv('~/data/baseline_prediction/stripped/geospatial.csv')
merged = merge(gf, geo_data, by='MRN')
# some variables are being read as numeric...
merged$Home_Price = as.numeric(merged$Home_Price)
merged$Fam_Income = as.numeric(merged$Fam_Income)
merged$Crime_Rate = as.numeric(merged$Crime_Rate)
phen_vars = c('SES', 'Home_Type', 'Home_Price', 'Fam_Income', 'Pop_BPL', 'Fam_BPL', 'Pub_School',
              'Crime_Rate', 'Green_Space', 'Park_Access', 'Air_Quality', 'Obesity_Rate',
              'Food_Index', 'Exercise_Access', 'Excessive_Drinking',
              'poverty_education', 'social_environment', 'physical_health_environment', 
              'physical_envronment')
keep_me = c()
for (v in phen_vars) {
  keep_me = c(keep_me, which(colnames(merged) == v))
}
X = merged[, keep_me]
geo_descriptive = cbind(nvVSadhd_test(X, merged, 'geo'),
                        nvVSadhd_lowVSadhd_high_test(X, merged),
                        outcome_test(X, merged),
                        symptom_test(X, merged, 0),
                        symptom_test(X, merged, 1),
                        symptom_test(X, merged, 2))
```

# Neuropsych
```{r}
beery_data = read.csv('~/data/baseline_prediction/stripped/beeryVMI.csv')
my_ids = intersect(gf$MRN, beery_data$Medical.Record...MRN)
mbeery = mergeOnClosestDate(gf, beery_data, my_ids, y.date='record.date.collected', y.id='Medical.Record...MRN')
rm_me = abs(mbeery$dateX.minus.dateY.months) > 12
mbeery = mbeery[!rm_me, ]
mbeery$dateClinical.minus.dateBeery.months = mbeery$dateX.minus.dateY.months
mbeery$dateX.minus.dateY.months = NULL

cpt_data = read.csv('~/data/baseline_prediction/stripped/cpt.csv')
my_ids = intersect(gf$MRN, cpt_data$MRN)
mcpt = mergeOnClosestDate(gf, cpt_data, my_ids)
rm_me = abs(mcpt$dateX.minus.dateY.months) > 12
mcpt = mcpt[!rm_me, ]
mcpt$dateClinical.minus.dateCPT.months = mcpt$dateX.minus.dateY.months
mcpt$dateX.minus.dateY.months = NULL

iq_data = read.csv('~/data/baseline_prediction/stripped/iq.csv')
my_ids = intersect(gf$MRN, iq_data$Medical.Record...MRN)
miq = mergeOnClosestDate(gf, iq_data, my_ids, y.id='Medical.Record...MRN', y.date='record.date.collected')
rm_me = abs(miq$dateX.minus.dateY.months) > 12
miq = miq[!rm_me, ]
miq$dateClinical.minus.dateIQ.months = miq$dateX.minus.dateY.months
miq$dateX.minus.dateY.months = NULL

wisc_data = read.csv('~/data/baseline_prediction/stripped/wisc.csv')
my_ids = intersect(gf$MRN, wisc_data$Medical.Record...MRN)
mwisc = mergeOnClosestDate(gf, wisc_data, my_ids, y.id='Medical.Record...MRN', y.date='record.date.collected')
rm_me = abs(mwisc$dateX.minus.dateY.months) > 12
mwisc = mwisc[!rm_me, ]
mwisc$dateClinical.minus.dateWISC.months = mwisc$dateX.minus.dateY.months
mwisc$dateX.minus.dateY.months = NULL

wj_data = read.csv('~/data/baseline_prediction/stripped/wj.csv')
my_ids = intersect(gf$MRN, wj_data$Medical.Record...MRN)
mwj = mergeOnClosestDate(gf, wj_data, my_ids, y.id='Medical.Record...MRN', y.date='record.date.collected')
rm_me = abs(mwj$dateX.minus.dateY.months) > 12
mwj = mwj[!rm_me, ]
mwj$dateClinical.minus.dateWJ.months = mwj$dateX.minus.dateY.months
mwj$dateX.minus.dateY.months = NULL

merged = merge(mwj, mbeery, by='MRN')
merged = merge(merged, miq, by='MRN')
merged = merge(merged, mwisc, by='MRN')
merged = merge(merged, mcpt, by='MRN')

phen_vars = c('FSIQ',
              # CPT
              'N_of_omissions', 'N_commissions', 'hit_RT', 'hit_RT_SE', 'variability_of_SE', 'N_perservations',
              'hit_RT_block_change', 'hit_RT_SE_block_change', 'hit_RT_ISI_change', 'hit_RT_SE_ISI_change',
              # WISC
              'Raw.score..DSF', 'Raw.score..DSB', 'Raw.score..SSF', 'Raw.score..SSB',
              # WJ
              'Raw.Score..VM', 'Raw.Score..DS', 'PS',
              # Beery
              'Standard.score'
              )
keep_me = c()
for (v in phen_vars) {
  keep_me = c(keep_me, which(colnames(merged) == v))
}
X = merged[, keep_me]
neuropsych_descriptive = cbind(nvVSadhd_test(X, merged, 'npsych'),
                        nvVSadhd_lowVSadhd_high_test(X, merged),
                        outcome_test(X, merged),
                        symptom_test(X, merged, 0),
                        symptom_test(X, merged, 1),
                        symptom_test(X, merged, 2))

```
# Structural
```{r}
struct_data = read.csv('~/data/baseline_prediction/stripped/structural.csv')
rm_me = (struct_data$mprage_score > 2)
struct_data = struct_data[!rm_me, ]
my_ids = intersect(gf$MRN, struct_data$MRN)
merged = mergeOnClosestDate(gf, struct_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
merged = merged[!rm_me, ]
X = merged[, 33:302]
struct_descriptive = cbind(nvVSadhd_test(X, merged, 'struct'),
                        nvVSadhd_lowVSadhd_high_test(X, merged),
                        outcome_test(X, merged),
                        symptom_test(X, merged, 0),
                        symptom_test(X, merged, 1),
                        symptom_test(X, merged, 2))
```

Finally, we just merge everything and spit out the file:

```{r}
res = rbind(dti_descriptive, demo_descriptive, geo_descriptive, neuropsych_descriptive,
            struct_descriptive)
write.csv(res, file='~/data/baseline_prediction/stripped/descriptives_age_sex.csv')
```

# Correlation with movement

Philip suggested we should check whether the DTI variables correlate with the movement variables. Let's see:

```{r}
tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
rm_me = (tract_data$fa_avg < .4 | tract_data$ad_avg < 1.18 | tract_data$rd_avg > .65 | tract_data$rd_avg < .5 |
         tract_data$norm.trans > .45 | tract_data$norm.rot > .008 | tract_data$goodSlices < 45 | tract_data$goodSlices > 70)
print(sprintf('Reducing from %d to %d scans', nrow(tract_data), nrow(tract_data)-sum(rm_me)))
tract_data = tract_data[!rm_me, ]
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf$MRN, tract_data$MRN)
merged = mergeOnClosestDate(gf, tract_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
print(sprintf('Reducing from %d to %d scans', nrow(merged), nrow(merged)-sum(rm_me)))
merged = merged[!rm_me, ]
phen_vars = c(which(grepl("^FA_", colnames(merged))),
              which(grepl("^AD_", colnames(merged))),
              which(grepl("^RD_", colnames(merged))),
              which(grepl("^MO_", colnames(merged)))
              )
X = merged[, phen_vars]
pvals.trans <- apply(X, 2, function(dat) {
  +      summary(lm(dat ~  merged$norm.trans))$coefficients[2, 4]})
pvals.rot <- apply(X, 2, function(dat) {
  +      summary(lm(dat ~  merged$norm.rot))$coefficients[2, 4]})
sum(pvals.trans < .05)
sum(pvals.rot < .05)
```

Well, 4 and 3 is not so bad, out of 44. Which ones?

```{r}
colnames(X)[pvals.trans < .05]
colnames(X)[pvals.rot < .05]
```

# Descriptives on deltas

Let's see which deltas (slopes) are correlated with the different targets. Let's do an uncorrected one, and then one correcting for sex, as age won't matter here.

```{r}
y_target = 'DX_BASELINE'

source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')

tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
rm_me = (tract_data$fa_avg < .4 | tract_data$ad_avg < 1.18 | tract_data$rd_avg > .65 | tract_data$rd_avg < .5 |
           tract_data$norm.trans > .45 | tract_data$norm.rot > .008 | tract_data$goodSlices < 45 |
           tract_data$goodSlices > 70)
tract_data = tract_data[!rm_me, ]
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)

phen_vars = c(which(grepl("^FA_", colnames(tract_data))),
              which(grepl("^AD_", colnames(tract_data))),
              which(grepl("^RD_", colnames(tract_data))),
              which(grepl("^MO_", colnames(tract_data)))
)
Xraw = tract_data[, phen_vars]
my_ids = unique(gf$MRN)

# MAKE SURE nrow(tract_data) == nrow(Xraw)!!!!

get_delta = function (d) {
  # if we have too many NAs, return NA
  if (sum(is.na(d)) >= (length(d)-1)) {
    return(NA)
  }
  else {
    lm(d ~ tract_data[idx, ]$age_at_scan)$coefficients[2]
  }
}

X = c()
y_idx = c()
target_col = which(colnames(gf)==y_target)
for (s in my_ids) {
  idx = tract_data$MRN==s
  # proceed if we have more than one observation in the data
  if (sum(idx) >= 2) {
    slopes = sapply(Xraw[idx, ], get_delta)
    names(slopes) = colnames(Xraw)
    X = rbind(X, slopes)
    idxy = gf$MRN==s
    y_idx = c(y_idx, which(idxy)[1])
  }
}
merged = gf[y_idx, ]
```

```{r}
dti_delta_descriptive = cbind(nvVSadhd_test(X, merged, 'DTI'),
                        nvVSadhd_lowVSadhd_high_test(X, merged),
                        outcome_test(X, merged))
```

Same thing for structural:

```{r}

```