---
title: "Prediction light"
output: html_notebook
---

I'll call this prediction light because I'm just using the variables I presented at the WIP. It might turn out that they're the best to go with anyways, but it's a limited set from what I've been working with. For example, we'll do no voxelwise stuff.

Let's start with a single seed (1234), then (1), then try several seeds to see what we get. Also, I'll start with a dataset that had very nice descriptive results (neuropsych), but also a few NAs if we use them all together. So, we'll need to either impute or use some algorithms that don't care about that.

```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf = gf[gf$BASELINE=='BASELINE' & gf$age <= 12, ]

beery_data = read.csv('~/data/baseline_prediction/stripped/beeryVMI.csv')
my_ids = intersect(gf$MRN, beery_data$Medical.Record...MRN)
mbeery = mergeOnClosestDate(gf, beery_data, my_ids, y.date='record.date.collected', y.id='Medical.Record...MRN')
rm_me = abs(mbeery$dateX.minus.dateY.months) > 12
mbeery = mbeery[!rm_me, ]
mbeery$dateClinical.minus.dateBeery.months = mbeery$dateX.minus.dateY.months
mbeery$dateX.minus.dateY.months = NULL

cpt_data = read.csv('~/data/baseline_prediction/stripped/cpt.csv')
my_ids = intersect(gf$MRN, cpt_data$MRN)
mcpt = mergeOnClosestDate(gf, cpt_data, my_ids)
rm_me = abs(mcpt$dateX.minus.dateY.months) > 12
mcpt = mcpt[!rm_me, ]
mcpt$dateClinical.minus.dateCPT.months = mcpt$dateX.minus.dateY.months
mcpt$dateX.minus.dateY.months = NULL

iq_data = read.csv('~/data/baseline_prediction/stripped/iq.csv')
my_ids = intersect(gf$MRN, iq_data$Medical.Record...MRN)
miq = mergeOnClosestDate(gf, iq_data, my_ids, y.id='Medical.Record...MRN', y.date='record.date.collected')
rm_me = abs(miq$dateX.minus.dateY.months) > 12
miq = miq[!rm_me, ]
miq$dateClinical.minus.dateIQ.months = miq$dateX.minus.dateY.months
miq$dateX.minus.dateY.months = NULL

wisc_data = read.csv('~/data/baseline_prediction/stripped/wisc.csv')
my_ids = intersect(gf$MRN, wisc_data$Medical.Record...MRN)
mwisc = mergeOnClosestDate(gf, wisc_data, my_ids, y.id='Medical.Record...MRN', y.date='record.date.collected')
rm_me = abs(mwisc$dateX.minus.dateY.months) > 12
mwisc = mwisc[!rm_me, ]
mwisc$dateClinical.minus.dateWISC.months = mwisc$dateX.minus.dateY.months
mwisc$dateX.minus.dateY.months = NULL

wj_data = read.csv('~/data/baseline_prediction/stripped/wj.csv')
my_ids = intersect(gf$MRN, wj_data$Medical.Record...MRN)
mwj = mergeOnClosestDate(gf, wj_data, my_ids, y.id='Medical.Record...MRN', y.date='record.date.collected')
rm_me = abs(mwj$dateX.minus.dateY.months) > 12
mwj = mwj[!rm_me, ]
mwj$dateClinical.minus.dateWJ.months = mwj$dateX.minus.dateY.months
mwj$dateX.minus.dateY.months = NULL

merged = merge(mwj, mbeery, by='MRN', all.x=T, all.y=T)
merged = merge(merged, miq, by='MRN', all.x=T, all.y=T)
merged = merge(merged, mwisc, by='MRN', all.x=T, all.y=T)
merged = merge(merged, mcpt, by='MRN', all.x=T, all.y=T)
# now age and SEX might have NAs if the entries came from two datasets, so let's grab
# them from gf again
age = c()
sex = c()
for (s in 1:nrow(merged)) {
  age = c(age, gf[gf$MRN==merged[s,]$MRN,]$age)
  sex = c(sex, gf[gf$MRN==merged[s,]$MRN,]$SEX)
}
SEX = factor(sex, labels = c('Female', 'Male'))
merged$age = NULL
merged$SEX = NULL
merged = cbind(merged, age, SEX)


phen_vars = c('FSIQ',
              # WISC
              'Raw.score..DSF', 'Raw.score..DSB', 'Raw.score..SSF', 'Raw.score..SSB',
              # WJ
              'Raw.Score..VM', 'Raw.Score..DS', 'PS',
              # Beery
              'Standard.score'
              )
keep_me = sapply(phen_vars, function(d) which(colnames(merged) == d))
X = merged[, keep_me]
```

*Note that since we're residualizing the entire dataset, I won't be a stickler for doing PCA and/or imputation in the entire data either. If we decide to not do that in the future, we can change it.*

```{r}
get_needed_residuals = function(y, fm_str, cutoff, df) {
  fm = as.formula(fm_str)
  fit = lm(fm)
  # selecting which covariates to use
  fm = "y ~ "
  for (r in 2:dim(summary(fit)$coefficients)[1]) {
    if (summary(fit)$coefficients[r, 4] < cutoff) {
      cname = rownames(summary(fit)$coefficients)[r]
      cname = gsub("SEXMale", "SEX", cname)
      fm = sprintf('%s + %s', fm, cname)
      print(fm)
    }
  }
  # don't do anything if no variables were significant
  if (fm != 'y ~ ') {
    opt_fit = lm(as.formula(fm))
    y[!is.na(y)] = opt_fit$residuals
  }
  return(y)
}

X_resid = sapply(X, get_needed_residuals, 'y ~ df$age + I(df$age^2) + df$SEX', .1, merged)
y = as.character(merged$DX_BASELINE)
y[y != 'NV'] = 'ADHD'
y = factor(y, levels=c('ADHD', 'NV'))
```

Now let's go for a simple classifier using a given random seed:

```{r}
myseed = 1234
tuneLength = 10
k = 5
times = 5
set.seed(myseed)
split <- createDataPartition(y, p = .8, list = FALSE)
Xtrain <- X_resid[ split, ]
ytrain <- y[ split ]
Xtest  <- X_resid[-split, ]
ytest = y[-split]

# library(parallel)
# cl <- makeCluster(8)
# pvals = parSapply(cl, Xtrain, function(d, ytrain) t.test(d ~ ytrain)$p.value, ytrain)
# stopCluster(cl)
# Xtrain = Xtrain[, which(pvals <= .05)]
# print(dim(Xtrain))
# 
# keep_me = sapply(colnames(Xtrain), function(d) which(colnames(Xtest) == d))
# Xtest = Xtest[, keep_me]
# 
# pp <- preProcess(Xtrain, method = c('BoxCox', 'center', 'scale', 'pca'), thresh=.9)
# filtXtrain<- predict(pp, Xtrain)
# filtXtest <- predict(pp, Xtest)
# print(dim(filtXtrain))

filtXtrain = Xtrain
filtXtest = Xtest

set.seed(myseed)
fullCtrl <- trainControl(method = "repeatedcv",
                         number = k,
                         repeats = times,
                         savePredictions="final",
                         classProbs=TRUE,
                         summaryFunction=twoClassSummary)

mymod <- train(filtXtrain, ytrain, tuneLength=tuneLength,
               trControl=fullCtrl, metric='ROC',
               method='ada')

```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).
