---
title: "Voxelwise slope baseline"
output: html_notebook
---

In both DTI and structural, let's use the voxelwise data to classify diagnostic at baseline. As usual, we start with the univariate filters, and then try eveyrthing.

# Thickness
```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
struct_data = read.csv('~/data/baseline_prediction/stripped/structural.csv')
load('~/data/baseline_prediction/struct_thickness.rData')
# the first column of lh and rh is an index variable
vdata = cbind(struct_data$Mask.ID...Scan,
              lh_thickness[,2:ncol(lh_thickness)],
              rh_thickness[,2:ncol(rh_thickness)])
rm(lh_thickness)
rm(rh_thickness)
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
```

But computing the slopes is taking a LONG time, so we should probably do it ahead of time, and just load it later. 
```{r}
get_delta = function (d, ages) {
  # if we have too many NAs, return NA
  if (sum(is.na(d)) >= (length(d)-1)) {
    return(NA)
  }
  else {
    lm(d ~ ages)$coefficients[2]
  }
}

X = c()
y = c()
target_col = which(colnames(gf_base)=='DX_BASELINE')
library(parallel)
cl <- makeCluster(32)
cnt = 0
for (s in unique(gf_base$MRN)) {
  print(sprintf('%d of %d', cnt + 1, length(unique(gf_base$MRN))))
  idx = struct_data$MRN==s
  # proceed if we have more than one observation in the data
  if (sum(idx) >= 2) {
    # slopes = sapply(vdata[idx, ], get_delta)
    slopes = parSapply(cl, vdata[idx, ], get_delta, struct_data[idx, ]$age_at_scan)
    X = rbind(X, slopes)
    idxy = gf_base$MRN==s
    y = c(y, unique(as.character(gf_base[idxy, target_col])))
  }
  cnt = cnt + 1
}
y[y!='NV'] = 'ADHD'
y = factor(y, levels = c('NV', 'ADHD'))
```
## Univariate filters

## Builtin feature selection

# Area
```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
struct_data = read.csv('~/data/baseline_prediction/stripped/structural.csv')
load('~/data/baseline_prediction/struct_area.rData')
# the first column of lh and rh is an index variable
vdata = cbind(struct_data$Mask.ID...Scan,
              lh_area[,2:ncol(lh_area)],
              rh_area[,2:ncol(rh_area)])
rm(lh_area)
rm(rh_area)
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
```

## Univariate filters

## Builtin feature selection

# Volume
```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
struct_data = read.csv('~/data/baseline_prediction/stripped/structural.csv')
load('~/data/baseline_prediction/struct_volume.rData')
# the first column of lh and rh is an index variable
vdata = cbind(volume_ids,
              lh_volume[,2:ncol(lh_volume)],
              rh_volume[,2:ncol(rh_volume)])
rm(lh_volume)
rm(rh_volume)
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
keep_me = c()
for (i in 1:nrow(vdata)) {
  if (vdata[i, 1] %in% struct_data$Mask.ID...Scan) {
    keep_me = c(keep_me, which(struct_data$Mask.ID...Scan == vdata[i, 1]))
  }
}
struct_data = struct_data[keep_me, ]
gf_base = gf[gf$BASELINE=='BASELINE', ]
```

## Univariate filters

## Builtin feature selection