---
title: "Analysis of structural data"
output: html_notebook
---

Now we do some classification using DX and structural data. First, some data cleaning:

```{r}
# for replication purposes
set.seed(107)
source('~/ncr_notebooks/baseline_prediction//src/aux_functions.R')

# Philip pre-selected which structural scans to use
gf_fname = '~/data/baseline_prediction/structural_match_02092017.csv'
data_fname = '~/data/baseline_prediction/structural_long_02092017.csv'
gf = read.csv(gf_fname)
group_var = 'DX2'
data = read.csv(data_fname)
maskid_gf = unique(gf[, c('MASKID', group_var)])

ldata = merge(maskid_gf, data, by.x='MASKID', by.y='lh.aparc.thickness', all.y = F)
eval(parse(text=sprintf('groups = ldata$\"%s\"', group_var)))
idx = groups=='NV' | groups=='ADHD'
groups = factor(groups[idx])
ldata = ldata[idx,]
```

As usual, let's start by throwing everything in there, and then start doing it by domains:

#All data

```{r}
phen_vars = c(which(grepl("_thickness$", colnames(ldata))),
              which(grepl("_volume$", colnames(ldata))),
              which(grepl("_area$", colnames(ldata))),
              which(grepl("^Left.", colnames(ldata))),
              which(grepl("^Right.", colnames(ldata)))
              )
# remove WM variables
idx = phen_vars < 261 | phen_vars > 266 
phen_vars = phen_vars[idx]
ldata = ldata[, phen_vars]

```

```{r}
root_fname = '~/data/baseline_prediction/results/structAll_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

#All data, PCA
```{r}
ldata.pca <- prcomp(ldata, center=T, scale=T)
plot(ldata.pca, type = "l")
print(ldata.pca$sdev^2)  # these are the eignevalues
```

This is interesting: if we go based on EV, we'd get the first 50. But based on elbow, it's only 3!

```{r}
ldata = ldata.pca$x[, 1:3]
root_fname = '~/data/baseline_prediction/results/structAll_PCAElbow_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

```{r}
ldata = ldata.pca$x[, 1:50]
root_fname = '~/data/baseline_prediction/results/structAll_PCAEV1_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

#Cortical thickness

```{r}
ldata = merge(maskid_gf, data, by.x='MASKID', by.y='lh.aparc.thickness', all.y = F)
eval(parse(text=sprintf('groups = ldata$\"%s\"', group_var)))
idx = groups=='NV' | groups=='ADHD'
groups = factor(groups[idx])
ldata = ldata[idx,]
phen_vars = which(grepl("_thickness$", colnames(ldata)))
# remove WM variables
idx = phen_vars < 261 | phen_vars > 266 
phen_vars = phen_vars[idx]
ldata = ldata[, phen_vars]

root_fname = '~/data/baseline_prediction/results/structThickness_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

#Cortical thickness, PCA

```{r}
ldata.pca <- prcomp(ldata, center=T, scale=T)
plot(ldata.pca, type = "l")
print(ldata.pca$sdev^2)  # these are the eignevalues
```
Same as before, EV1 gives us 14, but the elbow is only 2. 
```{r}
ldata = ldata.pca$x[, 1:2]
root_fname = '~/data/baseline_prediction/results/structThickness_PCAElbow_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

```{r}
ldata = ldata.pca$x[, 1:14]
root_fname = '~/data/baseline_prediction/results/structThickness_PCAEV1_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

#Cortical area

```{r}
ldata = merge(maskid_gf, data, by.x='MASKID', by.y='lh.aparc.thickness', all.y = F)
eval(parse(text=sprintf('groups = ldata$\"%s\"', group_var)))
idx = groups=='NV' | groups=='ADHD'
groups = factor(groups[idx])
ldata = ldata[idx,]
phen_vars = which(grepl("_area$", colnames(ldata)))
# remove WM variables
idx = phen_vars < 261 | phen_vars > 266 
phen_vars = phen_vars[idx]
ldata = ldata[, phen_vars]

root_fname = '~/data/baseline_prediction/results/structArea_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

#Cortical area, PCA

```{r}
ldata.pca <- prcomp(ldata, center=T, scale=T)
plot(ldata.pca, type = "l")
print(ldata.pca$sdev^2)  # these are the eignevalues
```
Same as before, EV1 gives us 14, but the elbow is only 2. 
```{r}
ldata = ldata.pca$x[, 1:2]
root_fname = '~/data/baseline_prediction/results/structArea_PCAElbow_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

```{r}
ldata = ldata.pca$x[, 1:14]
root_fname = '~/data/baseline_prediction/results/structArea_PCAEV1_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

#Cortical Volume

```{r}
ldata = merge(maskid_gf, data, by.x='MASKID', by.y='lh.aparc.thickness', all.y = F)
eval(parse(text=sprintf('groups = ldata$\"%s\"', group_var)))
idx = groups=='NV' | groups=='ADHD'
groups = factor(groups[idx])
ldata = ldata[idx,]
phen_vars = which(grepl("_volume$", colnames(ldata)))
# remove WM variables
idx = phen_vars < 261 | phen_vars > 266 
phen_vars = phen_vars[idx]
ldata = ldata[, phen_vars]

root_fname = '~/data/baseline_prediction/results/structVolume_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

#Cortical Volume, PCA

```{r}
ldata.pca <- prcomp(ldata, center=T, scale=T)
plot(ldata.pca, type = "l")
print(ldata.pca$sdev^2)  # these are the eignevalues
```
Same as before, EV1 gives us 16, but the elbow is only 2. 
```{r}
ldata = ldata.pca$x[, 1:2]
root_fname = '~/data/baseline_prediction/results/structVolume_PCAElbow_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```

```{r}
ldata = ldata.pca$x[, 1:16]
root_fname = '~/data/baseline_prediction/results/structVolume_PCAEV1_NVvsADHD'
if (! file.exists(sprintf('%s.RData', root_fname))) {
  save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
}
# ... waiting ...
res = collect_results(root_fname)
mylist = res[[1]]
run_models = res[[2]]
source('~/ncr_notebooks/baseline_prediction/src/do_metrics_plots.R')
```





