---
title: "Improving DTI results"
output: html_notebook
---

So, the goal is to improve on these results, either by using data from other domains, or better tuning the best model.

```{r}
library(pROC)
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
root_fname = '~/data/baseline_prediction/results/stripped_allDTI_NVvsADHD_big'
do_metrics_plots(root_fname)
```

Let's try better tunning of the RF model:

```{r}
myseed = 107
tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
rm_me = (tract_data$fa_avg < .4 | tract_data$ad_avg < 1.18 | tract_data$rd_avg > .65 | tract_data$rd_avg < .5 |
         tract_data$norm.trans > .45 | tract_data$norm.rot > .008 | tract_data$goodSlices < 45 | tract_data$goodSlices > 70)
print(sprintf('Reducing from %d to %d scans', nrow(tract_data), nrow(tract_data)-sum(rm_me)))
tract_data = tract_data[!rm_me, ]
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf$MRN, tract_data$MRN)
merged = mergeOnClosestDate(gf, tract_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
print(sprintf('Reducing from %d to %d scans', nrow(merged), nrow(merged)-sum(rm_me)))
merged = merged[!rm_me, ]

phen_vars = c(which(grepl("^FA_", colnames(merged))),
              which(grepl("^AD_", colnames(merged))),
              which(grepl("^RD_", colnames(merged))),
              which(grepl("^MO_", colnames(merged)))
              )
X = merged[, phen_vars]
y = as.character(merged$DX_BASELINE)
y[y != 'NV'] = 'ADHD'
y = factor(y, levels=c('NV', 'ADHD'))

# remove subjects with NA
rm_me = rowSums(is.na(X)) > 0
X = X[!rm_me, ]
y = y[!rm_me]

cpuDiff = 1
tuneLength = 10
default_preproc = c("center", 'scale')
set.seed(myseed)
split <- createDataPartition(y, p = .8, list = FALSE)

Xtrain <- X[ split, ]
ytrain <- y[ split ]
Xtest  <- X[-split, ]
ytest = y[-split]

set.seed(myseed)
index <- createMultiFolds(ytrain, k = 5, times = 5)

library(doMC)
ncpus <- detectBatchCPUs()
njobs <- ncpus - cpuDiff
registerDoMC(njobs)
selFunc = 'oneSE'  # maybe try oneSE and tolerance as well?

set.seed(myseed)
fullCtrl <- trainControl(method = "repeatedcv",
                         index = index,
                         search='grid',
                         summaryFunction = twoClassSummary,
                         classProbs = TRUE)
rndForestGrid <- expand.grid(.mtry=c(1:sqrt(ncol(Xtrain))))
rfFull <- train(Xtrain, ytrain,
                method = "rf",
                trControl = fullCtrl,
                tuneGrid=rndForestGrid,
                metric = 'ROC',
                preProcess = default_preproc)
  
rfFull
pred = predict(rfFull, Xtest)
postResample(pred, ytest)
roc(as.numeric(ytest), as.numeric(pred))

varSeq = 1:ncol(X)
ctrl <- rfeControl(method = "repeatedcv",
                   saveDetails = TRUE,
                   index = index,
                   returnResamp = "final")

ctrl$functions <- rfFuncs
set.seed(myseed)
rfRFE <- rfe(Xtrain, ytrain,
             sizes = varSeq,
             rfeControl = ctrl,
             preProcess = default_preproc,
             summaryFunction = twoClassSummary,
             metric = 'ROC')
rfRFE
pred = predict(rfRFE, Xtest)
postResample(pred, ytest)
roc(as.numeric(ytest), as.numeric(pred$pred))
```

It's worrisome to me that the ROC and accuracy results are not within the intervals we're seeing in the original plot. 

I just noticed that I won't be able to reproduce that result because I didn't set all the seeds I needed properly. 

Well, let's focus on the variables that have good T-stats for now. We will manually set them at first, but then we can do it within the CV loops:

```{r}
phen_vars = c('FA_left_unc', 'FA_right_cst')
keep_me = sapply(phen_vars, function(d) which(colnames(merged) == d))
X = merged[, keep_me]
y = as.character(merged$DX_BASELINE)
y[y != 'NV'] = 'ADHD'
y = factor(y, levels=c('NV', 'ADHD'))

# remove subjects with NA
rm_me = rowSums(is.na(X)) > 0
X = X[!rm_me, ]
y = y[!rm_me]

cpuDiff = 1
tuneLength = 10
default_preproc = c("center", 'scale')
set.seed(myseed)
split <- createDataPartition(y, p = .8, list = FALSE)

Xtrain <- X[ split, ]
ytrain <- y[ split ]
Xtest  <- X[-split, ]
ytest = y[-split]

set.seed(myseed)
index <- createMultiFolds(ytrain, k = 5, times = 5)

library(doMC)
ncpus <- detectBatchCPUs()
njobs <- ncpus - cpuDiff
registerDoMC(njobs)
selFunc = 'oneSE'  # maybe try oneSE and tolerance as well?

set.seed(myseed)
fullCtrl <- trainControl(method = "repeatedcv",
                         index = index,
                         search='grid',
                         summaryFunction = twoClassSummary,
                         classProbs = TRUE)
rndForestGrid <- expand.grid(.mtry=c(1:sqrt(ncol(Xtrain))))
rfFull <- train(Xtrain, ytrain,
                method = "rf",
                trControl = fullCtrl,
                tuneGrid=rndForestGrid,
                metric = 'ROC',
                preProcess = default_preproc)
  
rfFull
pred = predict(rfFull, Xtest)
postResample(pred, ytest)
roc(as.numeric(ytest), as.numeric(pred))

varSeq = 1:ncol(X)
ctrl <- rfeControl(method = "repeatedcv",
                   saveDetails = TRUE,
                   index = index,
                   returnResamp = "final")

ctrl$functions <- rfFuncs
set.seed(myseed)
rfRFE <- rfe(Xtrain, ytrain,
             sizes = varSeq,
             rfeControl = ctrl,
             preProcess = default_preproc,
             summaryFunction = twoClassSummary,
             metric = 'ROC')
rfRFE
pred = predict(rfRFE, Xtest)
postResample(pred, ytest)
roc(as.numeric(ytest), as.numeric(pred$pred))
```
