---
title: "Baseline descriptives"
output: html_notebook
---

As the first stage for the baseline paper, let's do the predictives to show what the data looks like in a univariate way. We will focus on baseline data predicting 4 different things:

* DX_baseline
* SX_baseline
* Latent class
* Annualized symptom change (possibly adjusted for baseline SX: Phili says dimensional outcome (final-initial/initial symptoms)/age diff)

And we can do that for every possible dataset we have. Shouldn't be that bad, as most of the scripts are done already. 

For the brain, let's do ROI/tract based, but also whole brain. Then we add the other datasets (geospatial, neuropsych, and eventually genetics). As usual, the most annoying step will be plotting everything, but hey-hoo.  

For now, let's focus on the classes first (i.e. DX and outcome), which are a bit easier and we don't have to worry about whether we threshold symptoms, or whether to include NVs in regressions. We can always to them later.

```{r}
nvVSadhd_test = function(pheno, df, fm_str) {
  # some struct variables are bad
  if (sum(pheno, na.rm=T) == 0) {
    tstat = NA
    pval = NA
  } else {
    if (! is.na(fm_str)) {
      cutoff = .1
      fm = as.formula(fm_str)
      
      fit = lm(fm)
      # selecting which covariates to use
      fm = "pheno ~ "
      # we do this before we add the y variable
      for (r in 2:nrow(summary(fit)$coefficients)) {
        if (summary(fit)$coefficients[r, 4] < cutoff) {
          cname = rownames(summary(fit)$coefficients)[r]
          cname = gsub("SEXMale", "SEX", cname)
          fm = sprintf('%s + %s', fm, cname)
        }
      }
      fm = sprintf('%s + y', fm)
    } else { fm = 'pheno ~ y' }
    
    print(fm)
    y = as.character(df$DX_BASELINE)
    y[y != 'NV'] = 'ADHD'
    y = factor(y, levels=c('ADHD', 'NV'))
    
    opt_fit = lm(as.formula(fm))
    mycoef = which(rownames(summary(opt_fit)$coefficients) == 'yNV')
    tstat <- summary(opt_fit)$coefficients[mycoef, 3]
    pval <- summary(opt_fit)$coefficients[mycoef, 4]
  }
  res = c(tstat, pval)
  names(res) = c('baseDX_tstat', 'baseDX_pval')
  return(res)
}

outcome_test = function(pheno, df, fm_str) {
  if (sum(pheno, na.rm = T) == 0) {
    fstat = NA
    pval = NA
    fstat2 = NA
    pval2 = NA
  } else {
    if (! is.na(fm_str)) {
      cutoff = .1
      # NOTE that in R, the covariates you want to remove in aov() should come first in the equation!
      fm = as.formula(fm_str)
      
      fit = lm(fm)
      # selecting which covariates to use
      fm = "pheno ~ "
      # we do this before we add the y variable
      for (r in 2:nrow(summary(fit)$coefficients)) {
        if (summary(fit)$coefficients[r, 4] < cutoff) {
          cname = rownames(summary(fit)$coefficients)[r]
          cname = gsub("SEXMale", "SEX", cname)
          fm = sprintf('%s + %s', fm, cname)
        }
      }
      fm = sprintf('%s + y', fm)
    }
    else { fm = 'pheno ~ y' }
    
    y = df$inatt3_named
    opt_fit = aov(lm(as.formula(fm)))
    mycoef = nrow(summary(opt_fit)[[1]]) - 1 # last row is residuals
    fstat <- summary(opt_fit)[[1]][mycoef, 4]
    pval <- summary(opt_fit)[[1]][mycoef, 5]
    
    y = df$HI3_named
    opt_fit = aov(lm(as.formula(fm)))
    mycoef = nrow(summary(opt_fit)[[1]]) - 1 # last row is residuals
    fstat2 <- summary(opt_fit)[[1]][mycoef, 4]
    pval2 <- summary(opt_fit)[[1]][mycoef, 5]
  }
  
  res = c(fstat, pval, fstat2, pval2)
  names(res) = c('inattOutcome_Fstat', 'inattOutcome_pval',
                 'hiOutcome_Fstat', 'hiOutcome_pval')
  return(res)
}

symptom_test = function(pheno, df, fm_str, mycoef) {
  fm = as.formula(fm_str)
  
  y = df$SX_inatt
  tstat <- summary(lm(fm))$coefficients[mycoef, 3]
  pval <- summary(lm(fm))$coefficients[mycoef, 4]
  
  y = df$SX_HI
  tstat2 <- summary(lm(fm))$coefficients[mycoef, 3]
  pval2 <- summary(lm(fm))$coefficients[mycoef, 4]
  
  res = c(tstat, pval, tstat2, pval2)
  names(res) = c('inattOutcome_tstat', 'inattOutcome_pval', 'hiOutcome_tstat', 'hiOutcome_pval')
  return(res)
}

# get_SX_slope = function (df) {
#   inatt = c()
#   hi = c()
#   mrns = c()
#   for (s in unique(df$MRN)) {
#     idx = df$MRN==s
#     # proceed if we have more than one observation in the data
#     if (sum(idx) >= 2) {
#       inatt = c(inatt, lm(SX_inatt ~ age_at_scan, data=df[idx,])$coefficients[2,1])
#       hi = c(hi, lm(SX_HI ~ age_at_scan, data=df[idx,])$coefficients[2, 1])
#       mrns = c(mrns, s)
#     }
#   }
#   res = cbind(inatt, hi)
#   colnames(res) = c('inatt', 'hi')
#   rownames(res) = mrns
#   return(res)
# }
```

# DTI
## Tract-based
```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
rm_me = (tract_data$fa_avg < .4 | tract_data$ad_avg < 1.18 | tract_data$rd_avg > .65 | tract_data$rd_avg < .5 |
         tract_data$norm.trans > .45 | tract_data$norm.rot > .008 | tract_data$goodSlices < 45 | tract_data$goodSlices > 70)
tract_data = tract_data[!rm_me, ]
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf_base$MRN, tract_data$MRN)
merged = mergeOnClosestDate(gf_base, tract_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
merged = merged[!rm_me, ]
phen_vars = c(which(grepl("^FA_", colnames(merged))),
              which(grepl("^AD_", colnames(merged))),
              which(grepl("^RD_", colnames(merged))),
              which(grepl("^MO_", colnames(merged)))
              )
X = merged[, phen_vars]
keep_me = merged$age <= 12
X = X[keep_me, ]
merged = merged[keep_me, ]
print(sprintf('DTI tracts: %d datapoints, %d phenotypes', nrow(X), ncol(X)))

res = sapply(X, nvVSadhd_test, merged, 'pheno ~ df$age + I(df$age^2) + df$SEX')
res2 = sapply(X, outcome_test, merged, 'pheno ~ df$age + I(df$age^2) + df$SEX')
write.csv(cbind(t(res), t(res2)), file='~/data/baseline_prediction/descriptives/dti_tracts.csv')
```
## Voxel-based

### FA
```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
load('~/data/baseline_prediction/dti/fa_voxelwise.RData')
dti_vdata = cbind(tract_data$maskid, fa_data)
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf_base$MRN, tract_data$MRN)
merged = mergeOnClosestDate(gf_base, tract_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
merged = merged[!rm_me, ]
dti_base_vdata = merge(merged$maskid, dti_vdata, by.x=1, by.y=1, all.y=F, all.x=T)

X = dti_base_vdata[, 2:ncol(dti_base_vdata)]
rm_me = colSums(is.na(X)) > 0
X = X[, !rm_me]
keep_me = merged$age <= 12
X = X[keep_me, ]
merged = merged[keep_me, ]
print(sprintf('FA voxels: %d datapoints, %d phenotypes', nrow(X), ncol(X)))
cl <- makeCluster(6)
res = parSapply(cl, X, nvVSadhd_test, merged, 'pheno ~ df$age + I(df$age^2) + df$SEX')
res2 = parSapply(cl, X, outcome_test, merged, 'pheno ~ df$age + I(df$age^2) + df$SEX')
stopCluster(cl)
```
```{r}
ijk = read.table('~/tmp/2021_fa.txt')
ijk[, 4] = res[1, ]
write.table(ijk, row.names=F, col.names=F, file=sprintf('~/data/baseline_prediction/descriptives/%s.txt',rownames(res)[1]))
ijk[, 4] = res[2, ]
write.table(ijk, row.names=F, col.names=F, file=sprintf('~/data/baseline_prediction/descriptives/%s.txt',rownames(res)[2]))
ijk[, 4] = p.adjust(res[2,], method='fdr')
write.table(ijk, row.names=F, col.names=F, file=sprintf('~/data/baseline_prediction/descriptives/%s_FDR.txt',rownames(res)[2]))

ijk[, 4] = res2[1, ]
write.table(ijk, row.names=F, col.names=F, file=sprintf('~/data/baseline_prediction/descriptives/%s.txt',rownames(res2)[1]))
ijk[, 4] = res2[2, ]
write.table(ijk, row.names=F, col.names=F, file=sprintf('~/data/baseline_prediction/descriptives/%s.txt',rownames(res2)[2]))
ijk[, 4] = p.adjust(res2[2,], method='fdr')
write.table(ijk, row.names=F, col.names=F, file=sprintf('~/data/baseline_prediction/descriptives/%s_FDR.txt',rownames(res2)[2]))
ijk[, 4] = res2[3, ]
write.table(ijk, row.names=F, col.names=F, file=sprintf('~/data/baseline_prediction/descriptives/%s.txt',rownames(res2)[3]))
ijk[, 4] = res2[4, ]
write.table(ijk, row.names=F, col.names=F, file=sprintf('~/data/baseline_prediction/descriptives/%s.txt',rownames(res2)[4]))
ijk[, 4] = p.adjust(res2[4,], method='fdr')
write.table(ijk, row.names=F, col.names=F, file=sprintf('~/data/baseline_prediction/descriptives/%s_FDR.txt',rownames(res2)[4]))
```
Then in bash:
```
m=RD
cat baseDX_tstat.txt | 3dUndump -master ~/data/baseline_prediction/dti/mean_fa_skeleton_mask.nii.gz -datum float -prefix tmp1 -overwrite -
cat baseDX_pval.txt | 3dUndump -master ~/data/baseline_prediction/dti/mean_fa_skeleton_mask.nii.gz -datum float -prefix tmp2 -overwrite -
cat baseDX_pval_FDR.txt | 3dUndump -master ~/data/baseline_prediction/dti/mean_fa_skeleton_mask.nii.gz -datum float -prefix tmp3 -overwrite -
3dbucket -prefix DTI_${m}_baseDX.nii tmp*
3drefit -sublabel 0 "tstat" -sublabel 1 "pval" -sublabel 2 "fdr" DTI_${m}_baseDX.nii 
rm tmp*

cat inattOutcome_Fstat.txt | 3dUndump -master ~/data/baseline_prediction/dti/mean_fa_skeleton_mask.nii.gz -datum float -prefix tmp1 -overwrite -
cat inattOutcome_pval.txt | 3dUndump -master ~/data/baseline_prediction/dti/mean_fa_skeleton_mask.nii.gz -datum float -prefix tmp2 -overwrite -
cat inattOutcome_pval_FDR.txt | 3dUndump -master ~/data/baseline_prediction/dti/mean_fa_skeleton_mask.nii.gz -datum float -prefix tmp3 -overwrite -
3dbucket -prefix DTI_${m}_inattOutcome.nii tmp*
3drefit -sublabel 0 "tstat" -sublabel 1 "pval" -sublabel 2 "fdr" DTI_${m}_inattOutcome.nii
rm tmp*

cat hiOutcome_Fstat.txt | 3dUndump -master ~/data/baseline_prediction/dti/mean_fa_skeleton_mask.nii.gz -datum float -prefix tmp1 -overwrite -
cat hiOutcome_pval.txt | 3dUndump -master ~/data/baseline_prediction/dti/mean_fa_skeleton_mask.nii.gz -datum float -prefix tmp2 -overwrite -
cat hiOutcome_pval_FDR.txt | 3dUndump -master ~/data/baseline_prediction/dti/mean_fa_skeleton_mask.nii.gz -datum float -prefix tmp3 -overwrite -
3dbucket -prefix DTI_${m}_hiOutcome.nii tmp*
3drefit -sublabel 0 "tstat" -sublabel 1 "pval" -sublabel 2 "fdr" DTI_${m}_hiOutcome.nii
rm tmp*

```
### AD
```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
load('~/data/baseline_prediction/dti/ad_voxelwise.RData')
dti_vdata = cbind(tract_data$maskid, ad_data)
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf_base$MRN, tract_data$MRN)
merged = mergeOnClosestDate(gf_base, tract_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
merged = merged[!rm_me, ]
dti_base_vdata = merge(merged$maskid, dti_vdata, by.x=1, by.y=1, all.y=F, all.x=T)

X = dti_base_vdata[, 2:ncol(dti_base_vdata)]
rm_me = colSums(is.na(X)) > 0
X = X[, !rm_me]
keep_me = merged$age <= 12
X = X[keep_me, ]
merged = merged[keep_me, ]
print(sprintf('AD voxels: %d datapoints, %d phenotypes', nrow(X), ncol(X)))
cl <- makeCluster(6)
res = parSapply(cl, X, nvVSadhd_test, merged, 'pheno ~ df$age + I(df$age^2) + df$SEX')
res2 = parSapply(cl, X, outcome_test, merged, 'pheno ~ df$age + I(df$age^2) + df$SEX')
stopCluster(cl)
```

### RD
```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
load('~/data/baseline_prediction/dti/rd_voxelwise.RData')
dti_vdata = cbind(tract_data$maskid, rd_data)
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf_base$MRN, tract_data$MRN)
merged = mergeOnClosestDate(gf_base, tract_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
merged = merged[!rm_me, ]
dti_base_vdata = merge(merged$maskid, dti_vdata, by.x=1, by.y=1, all.y=F, all.x=T)

X = dti_base_vdata[, 2:ncol(dti_base_vdata)]
rm_me = colSums(is.na(X)) > 0
X = X[, !rm_me]
keep_me = merged$age <= 12
X = X[keep_me, ]
merged = merged[keep_me, ]
print(sprintf('RD voxels: %d datapoints, %d phenotypes', nrow(X), ncol(X)))
cl <- makeCluster(6)
res = parSapply(cl, X, nvVSadhd_test, merged, 'pheno ~ df$age + I(df$age^2) + df$SEX')
res2 = parSapply(cl, X, outcome_test, merged, 'pheno ~ df$age + I(df$age^2) + df$SEX')
stopCluster(cl)
```

# Structural ROIs
```{r}
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
struct_data = read.csv('~/data/baseline_prediction/stripped/structural.csv')
rm_me = (struct_data$mprage_score > 2)
struct_data = struct_data[!rm_me, ]
my_ids = intersect(gf_base$MRN, struct_data$MRN)
merged = mergeOnClosestDate(gf_base, struct_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
merged = merged[!rm_me, ]
X = merged[, 33:302]
rm_me = colSums(is.na(X)) > 0
X = X[, !rm_me]
keep_me = merged$age <= 12
X = X[keep_me, ]
merged = merged[keep_me, ]
print(sprintf('Freesurfer ROIs: %d datapoints, %d phenotypes', nrow(X), ncol(X)))

cl <- makeCluster(6)
res = parSapply(cl, X, nvVSadhd_test, merged, 'pheno ~ df$age + I(df$age^2) + df$SEX')
res2 = parSapply(cl, X, outcome_test, merged, 'pheno ~ df$age + I(df$age^2) + df$SEX')
stopCluster(cl)
write.csv(cbind(t(res), t(res2)), file='~/data/baseline_prediction/descriptives/freesurfer_rois.csv')
```
# Structural voxelwise

## Area
```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
struct_data = read.csv('~/data/baseline_prediction/stripped/structural.csv')
load('~/data/baseline_prediction/struct_area.rData')
# the first column of lh and rh is an index variable
vdata = cbind(struct_data$Mask.ID...Scan,
              lh_area[,2:ncol(lh_area)],
              rh_area[,2:ncol(rh_area)])
rm(lh_area)
rm(rh_area)
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf_base$MRN, struct_data$MRN)
mstruct = mergeOnClosestDate(gf_base, struct_data, my_ids)
rm_me = abs(mstruct$dateX.minus.dateY.months) > 12
mstruct = mstruct[!rm_me, ]
struct_base_vdata = merge(mstruct$Mask.ID...Scan, vdata, by.x=1, by.y=1, all.y=F, all.x=T)
rm(vdata)
X = struct_base_vdata[, 2:ncol(struct_base_vdata)]
rm_me = colSums(is.na(X)) > 0
X = X[, !rm_me]
keep_me = mstruct$age <= 12
X = X[keep_me, ]
mstruct = mstruct[keep_me, ]
print(sprintf('Area voxels: %d datapoints, %d phenotypes', nrow(X), ncol(X)))
library(parallel)
cl <- makeCluster(6)
res = parSapply(cl, X, nvVSadhd_test, mstruct, 'pheno ~ df$age + I(df$age^2) + df$SEX')
res2 = parSapply(cl, X, outcome_test, mstruct, 'pheno ~ df$age + I(df$age^2) + df$SEX')
stopCluster(cl)
```

```{r}
res = rbind(res, p.adjust(res[2,], method='fdr'))
rownames(res)[3] = 'base_DX_pval_FDR'
res2 = rbind(res2, p.adjust(res2[2,], method='fdr'))
res2 = rbind(res2, p.adjust(res2[4,], method='fdr'))
rownames(res2)[5] = 'inattOutcome_pval_FDR'
rownames(res2)[6] = 'hiOutcome_pval_FDR'
write.csv(cbind(t(res), t(res2)), file='~/data/baseline_prediction/descriptives/thickness.csv', row.names=F)
```
## Thickness
```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
struct_data = read.csv('~/data/baseline_prediction/stripped/structural.csv')
load('~/data/baseline_prediction/struct_thickness.rData')
# the first column of lh and rh is an index variable
vdata = cbind(struct_data$Mask.ID...Scan,
              lh_thickness[,2:ncol(lh_thickness)],
              rh_thickness[,2:ncol(rh_thickness)])
rm(lh_thickness)
rm(rh_thickness)
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf_base$MRN, struct_data$MRN)
mstruct = mergeOnClosestDate(gf_base, struct_data, my_ids)
rm_me = abs(mstruct$dateX.minus.dateY.months) > 12
mstruct = mstruct[!rm_me, ]
struct_base_vdata = merge(mstruct$Mask.ID...Scan, vdata, by.x=1, by.y=1, all.y=F, all.x=T)
rm(vdata)
X = struct_base_vdata[, 2:ncol(struct_base_vdata)]
rm_me = colSums(is.na(X)) > 0
X = X[, !rm_me]
keep_me = mstruct$age <= 12
X = X[keep_me, ]
mstruct = mstruct[keep_me, ]
print(sprintf('Thickness voxels: %d datapoints, %d phenotypes', nrow(X), ncol(X)))
cl <- makeCluster(6)
res = parSapply(cl, X, nvVSadhd_test, mstruct, 'pheno ~ df$age + I(df$age^2) + df$SEX')
res2 = parSapply(cl, X, outcome_test, mstruct, 'pheno ~ df$age + I(df$age^2) + df$SEX')
stopCluster(cl)
```
## Volume
```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
struct_data = read.csv('~/data/baseline_prediction/stripped/structural.csv')
load('~/data/baseline_prediction/struct_volume.rData')
# the first column of lh and rh is an index variable
vdata = cbind(volume_ids,
              lh_volume[,2:ncol(lh_volume)],
              rh_volume[,2:ncol(rh_volume)])
rm(lh_volume)
rm(rh_volume)
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf_base$MRN, struct_data$MRN)
mstruct = mergeOnClosestDate(gf_base, struct_data, my_ids)
rm_me = abs(mstruct$dateX.minus.dateY.months) > 12
mstruct = mstruct[!rm_me, ]
keep_me = c()
keep_mstruct = c()
for (i in 1:nrow(vdata)) {
  if (vdata[i, 1] %in% mstruct$Mask.ID...Scan) {
    keep_me = c(keep_me, i)
    keep_mstruct = c(keep_mstruct, which(mstruct$Mask.ID...Scan == vdata[i, 1]))
  }
}
struct_base_vdata = vdata[keep_me, ]
mstruct = mstruct[keep_mstruct, ]
rm(vdata)

X = struct_base_vdata[, 2:ncol(struct_base_vdata)]
rm_me = colSums(is.na(X)) > 0
X = X[, !rm_me]
keep_me = mstruct$age <= 12
X = X[keep_me, ]
mstruct = mstruct[keep_me, ]
print(sprintf('Volume voxels: %d datapoints, %d phenotypes', nrow(X), ncol(X)))
cl <- makeCluster(6)
res = parSapply(cl, X, nvVSadhd_test, mstruct, 'pheno ~ df$age + I(df$age^2) + df$SEX')
res2 = parSapply(cl, X, outcome_test, mstruct, 'pheno ~ df$age + I(df$age^2) + df$SEX')
stopCluster(cl)
```

# Neuropsych
```{r}
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf = gf[gf$BASELINE=='BASELINE' & gf$age <= 12, ]

beery_data = read.csv('~/data/baseline_prediction/stripped/beeryVMI.csv')
my_ids = intersect(gf$MRN, beery_data$Medical.Record...MRN)
mbeery = mergeOnClosestDate(gf, beery_data, my_ids, y.date='record.date.collected', y.id='Medical.Record...MRN')
rm_me = abs(mbeery$dateX.minus.dateY.months) > 12
mbeery = mbeery[!rm_me, ]
mbeery$dateClinical.minus.dateBeery.months = mbeery$dateX.minus.dateY.months
mbeery$dateX.minus.dateY.months = NULL

cpt_data = read.csv('~/data/baseline_prediction/stripped/cpt.csv')
my_ids = intersect(gf$MRN, cpt_data$MRN)
mcpt = mergeOnClosestDate(gf, cpt_data, my_ids)
rm_me = abs(mcpt$dateX.minus.dateY.months) > 12
mcpt = mcpt[!rm_me, ]
mcpt$dateClinical.minus.dateCPT.months = mcpt$dateX.minus.dateY.months
mcpt$dateX.minus.dateY.months = NULL

iq_data = read.csv('~/data/baseline_prediction/stripped/iq.csv')
my_ids = intersect(gf$MRN, iq_data$Medical.Record...MRN)
miq = mergeOnClosestDate(gf, iq_data, my_ids, y.id='Medical.Record...MRN', y.date='record.date.collected')
rm_me = abs(miq$dateX.minus.dateY.months) > 12
miq = miq[!rm_me, ]
miq$dateClinical.minus.dateIQ.months = miq$dateX.minus.dateY.months
miq$dateX.minus.dateY.months = NULL

wisc_data = read.csv('~/data/baseline_prediction/stripped/wisc.csv')
my_ids = intersect(gf$MRN, wisc_data$Medical.Record...MRN)
mwisc = mergeOnClosestDate(gf, wisc_data, my_ids, y.id='Medical.Record...MRN', y.date='record.date.collected')
rm_me = abs(mwisc$dateX.minus.dateY.months) > 12
mwisc = mwisc[!rm_me, ]
mwisc$dateClinical.minus.dateWISC.months = mwisc$dateX.minus.dateY.months
mwisc$dateX.minus.dateY.months = NULL

wj_data = read.csv('~/data/baseline_prediction/stripped/wj.csv')
my_ids = intersect(gf$MRN, wj_data$Medical.Record...MRN)
mwj = mergeOnClosestDate(gf, wj_data, my_ids, y.id='Medical.Record...MRN', y.date='record.date.collected')
rm_me = abs(mwj$dateX.minus.dateY.months) > 12
mwj = mwj[!rm_me, ]
mwj$dateClinical.minus.dateWJ.months = mwj$dateX.minus.dateY.months
mwj$dateX.minus.dateY.months = NULL

res = nvVSadhd_test(miq$FSIQ, miq, NA)
res2 = outcome_test(miq$FSIQ, miq, NA)

phen_vars = c('N_of_omissions', 'N_commissions', 'hit_RT', 'hit_RT_SE', 'variability_of_SE', 'N_perservations',
              'hit_RT_block_change', 'hit_RT_SE_block_change', 'hit_RT_ISI_change', 'hit_RT_SE_ISI_change')
keep_me = sapply(phen_vars, function(d) which(colnames(mcpt) == d))
X = mcpt[, keep_me]
res = cbind(res, sapply(X, nvVSadhd_test, mcpt, 'pheno ~ df$age + df$SEX + I(df$age^2)'))
colnames(res)[1] = 'FSIQ'
res2 = cbind(res2, sapply(X, outcome_test, mcpt, 'pheno ~ df$age + df$SEX + I(df$age^2)'))
colnames(res2)[1] = 'FSIQ'

phen_vars = c('Raw.score..DSF', 'Raw.score..DSB', 'Raw.score..SSF', 'Raw.score..SSB')
keep_me = sapply(phen_vars, function(d) which(colnames(mwisc) == d))
X = mwisc[, keep_me]
res = cbind(res, sapply(X, nvVSadhd_test, mwisc, 'pheno ~ df$age + df$SEX + I(df$age^2)'))
res2 = cbind(res2, sapply(X, outcome_test, mwisc, 'pheno ~ df$age + df$SEX + I(df$age^2)'))

phen_vars = c('Raw.Score..VM', 'Raw.Score..DS')
keep_me = sapply(phen_vars, function(d) which(colnames(mwj) == d))
X = mwj[, keep_me]
res = cbind(res, sapply(X, nvVSadhd_test, mwj, 'pheno ~ df$age + df$SEX + I(df$age^2)'))
res2 = cbind(res2, sapply(X, outcome_test, mwj, 'pheno ~ df$age + df$SEX + I(df$age^2)'))

res = cbind(res, nvVSadhd_test(mwj$PS, mwj, NA))
res2 = cbind(res2, outcome_test(mwj$PS, mwj, NA))
colnames(res)[ncol(res)] = 'PS'
colnames(res2)[ncol(res2)] = 'PS'
res = cbind(res, nvVSadhd_test(mbeery$Standard.score, mbeery, NA))
res2 = cbind(res2, outcome_test(mbeery$Standard.score, mbeery, NA))
colnames(res)[ncol(res)] = 'Beery_VM'
colnames(res2)[ncol(res2)] = 'Beery_VM'

write.csv(cbind(t(res), t(res2)), file='~/data/baseline_prediction/descriptives/neuropsych.csv')
```


# Geospatial
```{r}
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
geo_data = read.csv('~/data/baseline_prediction/stripped/geospatial.csv')
merged = merge(gf_base, geo_data, by='MRN')
# some variables are being read as numeric...
merged$Home_Price = as.numeric(merged$Home_Price)
merged$Fam_Income = as.numeric(merged$Fam_Income)
merged$Crime_Rate = as.numeric(merged$Crime_Rate)
phen_vars = c('SES', 'Home_Type', 'Home_Price', 'Fam_Income', 'Pop_BPL', 'Fam_BPL', 'Pub_School',
              'Crime_Rate', 'Green_Space', 'Park_Access', 'Air_Quality', 'Obesity_Rate',
              'Food_Index', 'Exercise_Access', 'Excessive_Drinking',
              'poverty_education', 'social_environment', 'physical_health_environment', 
              'physical_envronment')
keep_me = c()
for (v in phen_vars) {
  keep_me = c(keep_me, which(colnames(merged) == v))
}
X = merged[, keep_me]
print(sprintf('Geospatial: %d datapoints, %d phenotypes', nrow(X), ncol(X)))
res = sapply(X, nvVSadhd_test, merged, 'pheno ~ df$age + df$SEX + I(df$age^2)')
res2 = sapply(X, outcome_test, merged, 'pheno ~ df$age + df$SEX +  + I(df$age^2)')
# write.csv(cbind(t(res), t(res2)), file='~/data/baseline_prediction/descriptives/geospatial.csv')
```
# Resting state

# Polygenic risk score
```{r}
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf_base = gf[gf$BASELINE=='BASELINE', ]
prs_data = read.csv('~/data/baseline_prediction/stripped/PRS.csv')
merged = merge(gf_base, prs_data, by='MRN')
X = merged[, 29:79]
print(sprintf('PRS: %d datapoints, %d phenotypes', nrow(X), ncol(X)))
res = sapply(X, nvVSadhd_test, merged, 'pheno ~ df$age + df$SEX + I(df$age^2)')
res2 = sapply(X, outcome_test, merged, 'pheno ~ df$age + df$SEX +  + I(df$age^2)')
# write.csv(cbind(t(res), t(res2)), file='~/data/baseline_prediction/descriptives/geospatial.csv')
```