---
title: "Differences in classification/regression results"
output: html_notebook
---

I was curious about the differences in using a regression, factor, or ordinal factor in classification. Would it make a difference? 

```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')

tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
rm_me = (tract_data$fa_avg < .4 | tract_data$ad_avg < 1.18 | tract_data$rd_avg > .65 | tract_data$rd_avg < .5 |
         tract_data$norm.trans > .45 | tract_data$norm.rot > .008 | tract_data$goodSlices < 45 | tract_data$goodSlices > 70)
print(sprintf('Reducing from %d to %d scans', nrow(tract_data), nrow(tract_data)-sum(rm_me)))
tract_data = tract_data[!rm_me, ]
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf$MRN, tract_data$MRN)
merged = mergeOnClosestDate(gf, tract_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
print(sprintf('Reducing from %d to %d scans', nrow(merged), nrow(merged)-sum(rm_me)))
merged = merged[!rm_me, ]

phen_vars = c(which(grepl("^FA_", colnames(merged))),
              which(grepl("^AD_", colnames(merged))),
              which(grepl("^RD_", colnames(merged))),
              which(grepl("^MO_", colnames(merged)))
              )
X = merged[, phen_vars]
y = merged$SX_inatt
rm_me = rowSums(is.na(X)) > 0
X = X[!rm_me, ]
merged = merged[!rm_me, ]
y = y[!rm_me]
summary(y)
class(y)
```

Now we do some very basic regression just do get a baseline:

```{r}
library(pROC)
myseed = 1234
cpuDiff = 1
tuneLength = 10
default_preproc = c("center", 'scale')
set.seed(myseed)
split <- createDataPartition(y, p = .8, list = FALSE)

Xtrain <- X[ split, ]
ytrain <- y[ split ]
Xtest  <- X[-split, ]
ytest = y[-split]

set.seed(myseed)
index <- createMultiFolds(ytrain, k = 5, times = 5)

library(doMC)
ncpus <- detectBatchCPUs()
njobs <- ncpus - cpuDiff
registerDoMC(njobs)
selFunc = 'best'  # maybe try oneSE and tolerance as well?

set.seed(myseed)
fullCtrl <- trainControl(method = "repeatedcv",
                         index = index,
                         search='grid')
rndForestGrid <- expand.grid(.mtry=c(1:sqrt(ncol(Xtrain))))
rfFull <- train(Xtrain, ytrain,
                method = "rf",
                trControl = fullCtrl,
                tuneGrid=rndForestGrid,
                preProcess = default_preproc)
  
getTrainPerf(rfFull)
pred = predict(rfFull, Xtest)
postResample(pred, ytest)
```
OK, now just to be safe, I'll run it again to show that we get the same results:
```{r}
set.seed(myseed)
rfFull2 <- train(Xtrain, ytrain,
                method = "rf",
                trControl = fullCtrl,
                tuneGrid=rndForestGrid,
                preProcess = default_preproc)
  
getTrainPerf(rfFull2)
pred = predict(rfFull2, Xtest)
postResample(pred, ytest)
```

All good. Now, we know that y is a continuous variable. What happens if we make it into a factor, thus running a classification now:

```{r}
y = factor(y)
Xtrain <- X[ split, ]
ytrain <- y[ split ]
Xtest  <- X[-split, ]
ytest = y[-split]

set.seed(myseed)
index <- createMultiFolds(ytrain, k = 5, times = 5)

set.seed(myseed)
fullCtrl <- trainControl(method = "repeatedcv",
                         index = index,
                         search='grid')
rndForestGrid <- expand.grid(.mtry=c(1:sqrt(ncol(Xtrain))))
rfFull <- train(Xtrain, ytrain,
                method = "rf",
                trControl = fullCtrl,
                tuneGrid=rndForestGrid,
                preProcess = default_preproc)
  
getTrainPerf(rfFull)
pred = predict(rfFull, Xtest)
postResample(pred, ytest)
```
We now switched to a 10-class classification, which we can reproduce here:
```{r}
set.seed(myseed)
rfFull2 <- train(Xtrain, ytrain,
                method = "rf",
                trControl = fullCtrl,
                tuneGrid=rndForestGrid,
                preProcess = default_preproc)
  
getTrainPerf(rfFull2)
pred = predict(rfFull2, Xtest)
postResample(pred, ytest)
```
The main question is whether I get different results using ordered factors. 
```{r}
y = factor(y, ordered=T)
head(y)
Xtrain <- X[ split, ]
ytrain <- y[ split ]
Xtest  <- X[-split, ]
ytest = y[-split]

set.seed(myseed)
index <- createMultiFolds(ytrain, k = 5, times = 5)

set.seed(myseed)
fullCtrl <- trainControl(method = "repeatedcv",
                         index = index,
                         search='grid')
rndForestGrid <- expand.grid(.mtry=c(1:sqrt(ncol(Xtrain))))
rfFull <- train(Xtrain, ytrain,
                method = "rf",
                trControl = fullCtrl,
                tuneGrid=rndForestGrid,
                preProcess = default_preproc)
  
getTrainPerf(rfFull)
pred = predict(rfFull, Xtest)
postResample(pred, ytest)
```
So, it looks like caret ignores that the factors are ordered, and uses them as regular classes.

Now, could we run a simple regression, and simply round the predictions between 0 and 9?

```{r}
X = merged[, phen_vars]
y = merged$SX_inatt
rm_me = rowSums(is.na(X)) > 0
X = X[!rm_me, ]
merged = merged[!rm_me, ]
y = y[!rm_me]
summary(y)
class(y)

set.seed(myseed)
split <- createDataPartition(y, p = .8, list = FALSE)
Xtrain <- X[ split, ]
ytrain <- y[ split ]
Xtest  <- X[-split, ]
ytest = y[-split]

set.seed(myseed)
index <- createMultiFolds(ytrain, k = 5, times = 5)

set.seed(myseed)
fullCtrl <- trainControl(method = "repeatedcv",
                         index = index,
                         search='grid')
rfFull <- train(Xtrain, ytrain,
                method = "rf",
                trControl = fullCtrl,
                tuneGrid=rndForestGrid,
                preProcess = default_preproc)
getTrainPerf(rfFull)
pred = predict(rfFull, Xtest)
```



