---
title: "DTI univariate selection"
output: html_notebook
---

Let's try some univariate selection with DTI.

```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')

tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
rm_me = (tract_data$fa_avg < .4 | tract_data$ad_avg < 1.18 | tract_data$rd_avg > .65 | tract_data$rd_avg < .5 |
           tract_data$norm.trans > .45 | tract_data$norm.rot > .008 | tract_data$goodSlices < 45 |
           tract_data$goodSlices > 70)
tract_data = tract_data[!rm_me, ]
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf$MRN, tract_data$MRN)
merged = mergeOnClosestDate(gf, tract_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
merged = merged[!rm_me, ]

phen_vars = c(which(grepl("^FA_", colnames(merged))),
              which(grepl("^AD_", colnames(merged))),
              which(grepl("^RD_", colnames(merged))),
              which(grepl("^MO_", colnames(merged)))
)
X = merged[, phen_vars]
y = merged$DX_BASELINE
y[y != 'NV'] = 'ADHD'
y = factor(y, levels = c('NV', 'ADHD'))
```

Let's get some baseline results before we start filtering:

```{r}
myseed = 1234
tuneLength = 10
cpuDiff = 1
mymod = 'rf'
library(pROC)
library(Hmisc)

set.seed(myseed)
split <- createDataPartition(y, p = .8, list = FALSE)
Xtrain <- X[ split, ]
ytrain <- y[ split ]
Xtest  <- X[-split, ]
ytest = y[-split]

pp = preProcess(Xtrain, method=c('YeoJohnson', 'center', 'scale', 'knnImpute'))
filtXtrain = predict(pp, Xtrain)
filtXtest = predict(pp, Xtest)

set.seed(myseed)
index <- createMultiFolds(ytrain, k = 5, times = 5)

library(doMC)
ncpus <- detectBatchCPUs()
njobs <- ncpus - cpuDiff
registerDoMC(njobs)
selFunc = 'oneSE'  # maybe try oneSE and tolerance as well?

set.seed(myseed)
fullCtrl <- trainControl(method = "repeatedcv",
                         index = index,
                         search='grid',
                         summaryFunction = twoClassSummary,
                         classProbs = TRUE)
m1 <- train(filtXtrain, ytrain,
            method = mymod,
            trControl = fullCtrl,
            tuneLength = tuneLength,
            metric = 'ROC')
print(getTrainPerf(m1))
pred = predict(m1, filtXtest)
print(postResample(pred, ytest))
print(roc(as.numeric(ytest), as.numeric(pred)))
```
These were very poor results, considering that our dumb classifier would get `r max(table(ytrain))/length(ytrain)`. If we add a 95% confidence interval, we're up to `r binconf(x=max(table(ytrain)), n=length(ytrain), alpha=.05)[3]`.

Now, let's use the pre-defined good variables, from the descriptives table:

```{r}
phen_vars = c('FA_right_slf',
              'FA_right_ifo','MO_right_ifo','FA_right_cst',
              'RD_right_slf',
              'FA_right_ilf',
              'FA_left_cst',
              'MO_left_ifo',
              'FA_left_ifo',
              'RD_right_ifo',
              'FA_right_unc',
              'MO_right_ilf',
              'MO_right_slf',
              'RD_right_cst','RD_right_ilf','MO_left_ilf','FA_cc',
              'MO_left_slf','FA_left_unc')
keep_me = sapply(phen_vars, function(d) which(colnames(merged) == d))
X = merged[, keep_me]
y = merged$DX_BASELINE
y[y != 'NV'] = 'ADHD'
y = factor(y, levels = c('NV', 'ADHD'))

Xtrain <- X[ split, ]
ytrain <- y[ split ]
Xtest  <- X[-split, ]
ytest = y[-split]

pp = preProcess(Xtrain, method=c('YeoJohnson', 'center', 'scale', 'knnImpute'))
filtXtrain = predict(pp, Xtrain)
filtXtest = predict(pp, Xtest)

m1 <- train(filtXtrain, ytrain,
            method = mymod,
            trControl = fullCtrl,
            tuneLength = tuneLength,
            metric = 'ROC')
print(getTrainPerf(m1))
pred = predict(m1, filtXtest)
print(postResample(pred, ytest))
print(roc(as.numeric(ytest), as.numeric(pred)))
```





