---
title: "Age restricted DTI analysis using stripped files"
output: html_notebook
---

Let's first plot the DTI results we're comparing to, before we do the age restriction analysis:

```{r}
source('~/ncr_notebooks/baseline_prediction/src/aux_functions.R')
root_fname = '~/data/baseline_prediction/results/stripped_allDTI_NVvsADHD_big'
do_metrics_plots(root_fname)
```

Now we can run a series of age restrictions to the data, and see if it gets better or worse:

```{r}
tract_data = read.csv('~/data/baseline_prediction/stripped/dti.csv')
rm_me = (tract_data$fa_avg < .4 | tract_data$ad_avg < 1.18 | tract_data$rd_avg > .65 | tract_data$rd_avg < .5 |
         tract_data$norm.trans > .45 | tract_data$norm.rot > .008 | tract_data$goodSlices < 45 | tract_data$goodSlices > 70)
print(sprintf('Reducing from %d to %d scans', nrow(tract_data), nrow(tract_data)-sum(rm_me)))
tract_data = tract_data[!rm_me, ]
gf_fname = '~/data/baseline_prediction/stripped/clinical.csv'
gf = read.csv(gf_fname)
gf = gf[gf$BASELINE=='BASELINE', ]
my_ids = intersect(gf$MRN, tract_data$MRN)
merged = mergeOnClosestDate(gf, tract_data, my_ids)
rm_me = abs(merged$dateX.minus.dateY.months) > 12
print(sprintf('Reducing from %d to %d scans', nrow(merged), nrow(merged)-sum(rm_me)))
merged = merged[!rm_me, ]
```

```{r}
library(caret)
library(randomForest)
seed = 107
phen_vars = c(which(grepl("^FA_", colnames(merged))),
              which(grepl("^AD_", colnames(merged))),
              which(grepl("^RD_", colnames(merged))),
              which(grepl("^MO_", colnames(merged)))
              )
X = merged[, phen_vars]
y = merged$DX_BASELINE
y[y != 'NV'] = 'ADHD'
y = factor(y)
```

```{r}
hist(merged$age_at_scan, breaks=40)
```

Just based on the histogram, let's start cutting off anyone above 13, then we go down to assess a few more theory-inspired cutoffs:

```{r}
# rm_me = (merged$age_at_scan > 13) | (rowSums(is.na(X)) > 0)
# ldata = X[!rm_me, ]
# groups = y[!rm_me]
root_fname = '~/data/baseline_prediction/results/stripped_allDTISE13_NVvsADHD'
# save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)

# #runInCluster(root_fname, train_test_ratio=.8, cpuDiff=0, nsplits=30, nrepeatcv=5, run_models=c('rndForest', 'lr', 'xgb'))
# 
do_metrics_plots(root_fname)
```

```{r}
# rm_me = (merged$age_at_scan > 12) | (rowSums(is.na(X)) > 0)
# ldata = X[!rm_me, ]
# groups = y[!rm_me]
root_fname = '~/data/baseline_prediction/results/stripped_allDTISE12_NVvsADHD'
# save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)

# #runInCluster(root_fname, train_test_ratio=.8, cpuDiff=0, nsplits=30, nrepeatcv=5, run_models=c('rndForest', 'lr', 'xgb'))
# 
do_metrics_plots(root_fname)
```

```{r}
# rm_me = (merged$age_at_scan > 11) | (rowSums(is.na(X)) > 0)
# ldata = X[!rm_me, ]
# groups = y[!rm_me]
root_fname = '~/data/baseline_prediction/results/stripped_allDTISE11_NVvsADHD'
# save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)

#runInCluster(root_fname, train_test_ratio=.8, cpuDiff=0, nsplits=30, nrepeatcv=5, run_models=c('rndForest', 'lr', 'xgb'))

do_metrics_plots(root_fname)
```

```{r}
# rm_me = (merged$age_at_scan > 10) | (rowSums(is.na(X)) > 0)
# ldata = X[!rm_me, ]
# groups = y[!rm_me]
root_fname = '~/data/baseline_prediction/results/stripped_allDTISE10_NVvsADHD'
# save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)

#runInCluster(root_fname, train_test_ratio=.8, cpuDiff=0, nsplits=30, nrepeatcv=5, run_models=c('rndForest', 'lr', 'xgb'))

do_metrics_plots(root_fname)
```

```{r}
# rm_me = (merged$age_at_scan > 9) | (rowSums(is.na(X)) > 0)
# ldata = X[!rm_me, ]
# groups = y[!rm_me]
root_fname = '~/data/baseline_prediction/results/stripped_allDTISE9_NVvsADHD'
# save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)
# 
#runInCluster(root_fname, train_test_ratio=.8, cpuDiff=0, nsplits=30, nrepeatcv=5, run_models=c('rndForest', 'lr', 'xgb'))
do_metrics_plots(root_fname)
```

```{r}
# rm_me = (merged$age_at_scan > 8) | (rowSums(is.na(X)) > 0)
# ldata = X[!rm_me, ]
# groups = y[!rm_me]
root_fname = '~/data/baseline_prediction/results/stripped_allDTISE8_NVvsADHD'
# save(ldata, groups, file=sprintf('%s.RData', root_fname), compress=T)

#runInCluster(root_fname, train_test_ratio=.8, cpuDiff=0, nsplits=30, nrepeatcv=5, run_models=c('rndForest', 'lr', 'xgb'))

do_metrics_plots(root_fname)
```

The results are a bit concerning. It looks like we need all that data, in particular the data from the older kids, in order to make a baseline prediction. Not sure how clinically informative that is. Should we try to really fit our models within the < 8 y.o range, and then expand it?

The trick is that I only have 76 scans < 8, but 120 < 9. That's a more manageable number that we could work with in better fitting classifiers.

Another way of getting more scans would be to possibly derive the visual thresholds from histograms in the age subgroups. Maybe that would savage  some of the data?