---
title: "Reviewing ICA method"
output: html_notebook
---

This is the comment:

_3. I would also like the authors to comment on the possibility of circular inference in the following procedure they applied: adding the mean effect of inattention/hyperactivity symptoms to the connectivity matrices, performing ICASSO on them and then applying the stable components on the data to determine whether those components are associated with the severity of the symptoms._

We will have to address the main theoretical comment later. But in terms of additional analysis, we can ask the following:

*Does ICASSO favor components associated with SX?*

Before we go ahead, let's export just the IQs from MATLAB into a CSV:

```
load('/data/NCR_SBRB/ica_results_enrichedNoDummies1K.mat')
dlmwrite('~/tmp/fmri_iqs.csv,iq,',')
```

We can run the inattention and HI regressions for all ICs, and sort them based on IQ.

```{r}
fmri_iqs = read.csv('~/data/fmri/fmri_iqs.csv', header = F)[,1]
library(RcppCNPy)
fmat <- npyLoad('/Users/sudregp/data/sripada/fmri/exp_scores_enrichedNoDummies1K.npy')
subjs = read.table('~/data/fmri/joel_all_147.txt')
exp_scores = cbind(subjs, fmat)
gf = read.csv('~/data/fmri/gf.csv')
df = merge(gf, exp_scores, by.x='maskid', by.y='V1')
df = df[df$used==1,]
adhdOnly = df$group!='NV'
pvals_inatt = sapply(df[adhdOnly, 14:ncol(df)], function(d) cor.test(d, df[adhdOnly,]$inatt, method='spearman')$p.value)
pvals_hi = sapply(df[adhdOnly, 14:ncol(df)], function(d) cor.test(d, df[adhdOnly,]$hi, method='spearman')$p.value)
```

Now we make the fMRI plot:

```{r}
myorder = sort(fmri_iqs, index.return=T, decreasing = T)
par(mfrow=(c(1, 2)))
plot(fmri_iqs[myorder$ix], pvals_inatt[myorder$ix], xlab='Stability', ylab='P-value', main='Inattention')
abline(h=.05, lty=3)
abline(v=.85, lty=3)
plot(fmri_iqs[myorder$ix], pvals_hi[myorder$ix], xlab='Stability', ylab='P-value', main='HI')
abline(h=.05, lty=3)
abline(v=.85, lty=3)
cor.test(1-pvals_inatt, fmri_iqs, method='spearman')
cor.test(1-pvals_hi, fmri_iqs, method='spearman')
```

So, there's clearly no relationship between symptom count and how stable the IC actually is. Let's try it in MEG:

```{r}
bands = c('1-4', '4-8', '8-13', '13-30', '30-55', '65-100')
meg_iqs = c()
fmat = c()
for (b in bands) {
  meg_iqs = c(meg_iqs, read.csv(sprintf('~/data/meg/iqs_%s.csv', b), header = F)[,1])
  fmat = cbind(fmat, npyLoad(sprintf('/Users/sudregp/data/sripada/meg/exp_scores_%s_imcoh_enriched_1000perms_15ics.npy', b)))
}
subjs = read.table('~/data/meg/usable_subjects_5segs13p654_SAM.txt')
exp_scores = cbind(subjs, fmat)
gf = read.csv('~/data/meg/gf.csv')
df = merge(gf, exp_scores, by.x='maskid', by.y='V1')
df = df[df$used==1,]
adhdOnly = df$group!='NV'
pvals_inatt = sapply(df[adhdOnly, 17:ncol(df)], function(d) cor.test(d, df[adhdOnly,]$inatt, method='spearman')$p.value)
pvals_hi = sapply(df[adhdOnly, 17:ncol(df)], function(d) cor.test(d, df[adhdOnly,]$hi, method='spearman')$p.value)
```

Let's look at the MEG plots:

```{r}
myorder = sort(meg_iqs, index.return=T, decreasing = T)
par(mfrow=(c(1, 2)))
plot(meg_iqs[myorder$ix], pvals_inatt[myorder$ix], xlab='Stability', ylab='P-value', main='Inattention')
abline(h=.05, lty=3)
abline(v=.85, lty=3)
plot(meg_iqs[myorder$ix], pvals_hi[myorder$ix], xlab='Stability', ylab='P-value', main='HI')
abline(h=.05, lty=3)
abline(v=.85, lty=3)
cor.test(1-pvals_inatt, meg_iqs, method='spearman')
cor.test(1-pvals_hi, meg_iqs, method='spearman')
```

Again, it doesn't look like a more stable IC is more likely to be significantly correlated to SX.

# Theoretical component

This "augmented" ICA cannot be considered to be an unsupervised method. Regular ICAs, PCAs, or any other unsupervised dimensionality reduction methods do not take into consideration group membership or other possibly related variables. Here, we purposely used the augmented ICA method described in (Sripada REF) to tease out independent components that are also related to the total symptom count. 

As stated in (Sripada REF), prior to the ICA analysis, the regression-based cleaning procedure of each individual connection (regression targets) involves using both the covariates of interest (i.e. inattention and HI symptoms) and nuisance covariates (i.e. sex, age, movement, movement squared) as predictors so that weak collinearities between effects of interest and nuisance effects do not result in removal of interesting signal. We then augment each (cleaned) connection by its mean effect for features of interest to induce selection of components related to these phenotypes of interest. Note here that we are effectively adding a constant for each connection, prior to running ICA.

We finally use nonparametric tests (corrected using Bonferroni) to tease out which stable components are independently related to inattention and HI symptoms. ICASSO (as illustrated above) shows no significant bias in attaching a higher stability metric to ICs more correlated to symptom counts. Also note that the majority of components (stable or not) show no relationship to the two phenotypes of interest.

# Correlation to total symptoms

But it so happens that we augmented the ICA with total symptoms, and not inattention or HI by itself. So, let's see if the same analysis as above works for total symptoms.

```{r}
fmri_iqs = read.csv('~/data/fmri/fmri_iqs.csv', header = F)[,1]
library(RcppCNPy)
fmat <- npyLoad('/Users/sudregp/data/sripada/fmri/exp_scores_enrichedNoDummies1K.npy')
subjs = read.table('~/data/fmri/joel_all_147.txt')
exp_scores = cbind(subjs, fmat)
gf = read.csv('~/data/fmri/gf.csv')
df = merge(gf, exp_scores, by.x='maskid', by.y='V1')
df = df[df$used==1,]
total = df$inatt + df$hi
pvals = sapply(df[, 14:ncol(df)], function(d) cor.test(d, total, method='spearman')$p.value)
adhdOnly = df$group!='NV'
total = df[adhdOnly,]$inatt + df[adhdOnly,]$hi
pvals_adhdOnly = sapply(df[adhdOnly, 14:ncol(df)], function(d) cor.test(d, total, method='spearman')$p.value)
myorder = sort(fmri_iqs, index.return=T, decreasing = T)
par(mfrow=(c(1, 2)))
plot(fmri_iqs[myorder$ix], pvals[myorder$ix], xlab='Stability', ylab='P-value', main='Everyone')
abline(h=.05, lty=3)
abline(v=.85, lty=3)
plot(fmri_iqs[myorder$ix], pvals_adhdOnly[myorder$ix], xlab='Stability', ylab='P-value', main='ADHD Only')
abline(h=.05, lty=3)
abline(v=.85, lty=3)
cor.test(1-pvals, fmri_iqs, method='spearman', alternative='greater')
cor.test(1-pvals_adhdOnly, fmri_iqs, method='spearman', alternative='greater')
```


Still nothing for fMRI. Let's check on MEG:

```{r}
bands = c('1-4', '4-8', '8-13', '13-30', '30-55', '65-100')
meg_iqs = c()
fmat = c()
for (b in bands) {
  meg_iqs = c(meg_iqs, read.csv(sprintf('~/data/meg/iqs_%s.csv', b), header = F)[,1])
  fmat = cbind(fmat, npyLoad(sprintf('/Users/sudregp/data/sripada/meg/exp_scores_%s_imcoh_enriched_1000perms_15ics.npy', b)))
}
subjs = read.table('~/data/meg/usable_subjects_5segs13p654_SAM.txt')
exp_scores = cbind(subjs, fmat)
gf = read.csv('~/data/meg/gf.csv')
df = merge(gf, exp_scores, by.x='maskid', by.y='V1')
df = df[df$used==1,]
total = df$inatt + df$hi
pvals = sapply(df[, 17:ncol(df)], function(d) cor.test(d, total, method='spearman')$p.value)
adhdOnly = df$group!='NV'
total = df[adhdOnly,]$inatt + df[adhdOnly,]$hi
pvals_adhdOnly = sapply(df[adhdOnly, 17:ncol(df)], function(d) cor.test(d, total, method='spearman')$p.value)
myorder = sort(meg_iqs, index.return=T, decreasing = T)
par(mfrow=(c(1, 2)))
plot(meg_iqs[myorder$ix], pvals[myorder$ix], xlab='Stability', ylab='P-value', main='All')
abline(h=.05, lty=3)
abline(v=.85, lty=3)
plot(meg_iqs[myorder$ix], pvals_adhdOnly[myorder$ix], xlab='Stability', ylab='P-value', main='ADHD only')
abline(h=.05, lty=3)
abline(v=.85, lty=3)
cor.test(1-pvals, meg_iqs, method='spearman', alternative='greater')
cor.test(1-pvals_adhdOnly, meg_iqs, method='spearman', alternative='greater')
```

# Association with total symptoms

```{r}
fmri_iqs = read.csv('~/data/fmri/fmri_iqs.csv', header = F)[,1]
library(RcppCNPy)
fmat <- npyLoad('/Users/sudregp/data/sripada/fmri/exp_scores_enrichedNoDummies1K.npy')
subjs = read.table('~/data/fmri/joel_all_147.txt')
exp_scores = cbind(subjs, fmat)
gf = read.csv('~/data/fmri/gf.csv')
df = merge(gf, exp_scores, by.x='maskid', by.y='V1')
df = df[df$used==1,]
total = df$inatt + df$hi
pvals = sapply(df[, 14:ncol(df)], function(d) cor.test(d, total, method='spearman')$p.value)
adhdOnly = df$group!='NV'
total = df[adhdOnly,]$inatt + df[adhdOnly,]$hi
pvals_adhdOnly = sapply(df[adhdOnly, 14:ncol(df)], function(d) cor.test(d, total, method='spearman')$p.value)
sum(pvals_adhdOnly[fmri_iqs > .85] < .05/6)

bands = c('1-4', '4-8', '8-13', '13-30', '30-55', '65-100')
meg_iqs = c()
fmat = c()
for (b in bands) {
  meg_iqs = c(meg_iqs, read.csv(sprintf('~/data/meg/iqs_%s.csv', b), header = F)[,1])
  fmat = cbind(fmat, npyLoad(sprintf('/Users/sudregp/data/sripada/meg/exp_scores_%s_imcoh_enriched_1000perms_15ics.npy', b)))
}
subjs = read.table('~/data/meg/usable_subjects_5segs13p654_SAM.txt')
exp_scores = cbind(subjs, fmat)
gf = read.csv('~/data/meg/gf.csv')
df = merge(gf, exp_scores, by.x='maskid', by.y='V1')
df = df[df$used==1,]
total = df$inatt + df$hi
pvals = sapply(df[, 17:ncol(df)], function(d) cor.test(d, total, method='spearman')$p.value)
adhdOnly = df$group!='NV'
total = df[adhdOnly,]$inatt + df[adhdOnly,]$hi
pvals_adhdOnly = sapply(df[adhdOnly, 17:ncol(df)], function(d) cor.test(d, total, method='spearman')$p.value)
sum(pvals_adhdOnly[meg_iqs > .85] < .05/56)
```


