---
title: "Using QC and age groups"
output: html_notebook
---

Let's see what happens, especially with DTI, as we restrict by QC and age groups:

```{r}
gf = read.csv('~/data/prs/clinical_06192017.csv')
gf = gf[gf$ADHD_current_yes_no!='exclude',]
pgc = read.csv('~/data/prs/PRS2017_original_clump_default.csv')
df = merge(gf, pgc)
# remove duplicated MRNs
df = df[!duplicated(df$MRN),]
```

```{r}
library(mediation)
run_mediation = function(data, X, M, Y, nboot=100, short=T) {
    data = data[!is.na(data[, M]), ]
    data[, c(X, M)] = scale(data[, c(X, M)])
    data = as.data.frame(data)
    
    fm = as.formula(sprintf('%s ~ %s', M, X))
    fy = as.formula(sprintf('%s ~ %s + %s', Y, X, M))
    model.M <- lm(fm, data=data)
    model.Y <- glm(fy, data=data, family=binomial(link='logit'))
    results <- mediate(model.M, model.Y, treat=X, mediator=M, boot=T, sims=nboot, boot.ci.type='bca')
    if (short) {
      res = c(results$mediator, results$nobs, results$tau.coef, results$tau.p, results$d.avg, results$d.avg.p,
              results$z.avg, results$z.avg.p, results$n.avg, results$n.avg.p)
      names(res) = c('M', 'nobs', 'tot', 'tot_p', 'acme', 'acme_p', 'ade', 'ade_p', 'prop', 'prop_p')
      return(res)
    } else {
      return(results)
    }
}
```

Let's see make a few plots to decide on thresholds:

```{r}
dti = read.csv('~/data/prs/dti_07062017.csv')
tracts = merge(df, dti, by='MRN')
par(mfrow=c(3, 3))
hist(tracts$age, breaks=20)
hist(tracts$numVolumes, breaks=20)
hist(tracts$norm.rot, breaks=20)
hist(tracts$norm.trans, breaks=20)
hist(tracts$mean_fa, breaks=20)
hist(tracts$mean_ad, breaks=20)
hist(tracts$mean_rd, breaks=20)
```

Let's look at all DTI again, but without the bad data based on histograms:

```{r}
rm_me = tracts$numVolumes < 60 | tracts$norm.rot > .003 | tracts$norm.trans > .3 | tracts$mean_fa < .3 | tracts$mean_ad < .97 | tracts$mean_rd < .5
dti_tracts = tracts[!rm_me, ]

res = c()
for (m in colnames(dti_tracts)[56:ncol(dti_tracts)]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(dti_tracts, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
write.csv(res, file='~/data/prs/results_PRSp3_DTI_DX_histogramCleaned.csv', row.names=F)
```

Now, let's make it even more stringent, restricting noise within phenotype:

```{r}
for (j in 56:ncol(dti_tracts)) {
  upper = mean(dti_tracts[, j], na.rm=T) + 2*sd(dti_tracts[, j], na.rm=T)
  lower = mean(dti_tracts[, j], na.rm=T) - 2*sd(dti_tracts[, j], na.rm=T)
  idx = which(dti_tracts[, j] < lower | dti_tracts[, j] > upper)
  dti_tracts[idx, j] = NA
}
res = c()
for (m in colnames(dti_tracts)[56:ncol(dti_tracts)]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(dti_tracts, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
write.csv(res, file='~/data/prs/results_PRSp3_DTI_DX_histogramCleanedWithinTract2SD.csv', row.names=F)
```

Finally, let's see how this looks just for restricted age ranges:

```{r}
res = c()
for (m in colnames(dti_tracts)[56:ncol(dti_tracts)]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(dti_tracts[dti_tracts$age_at_scan < 21, ], 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
write.csv(res, file='~/data/prs/results_PRSp3_DTI_DX_histogramCleanedWithinTract2SD_AgeSE21.csv', row.names=F)
```

```{r}
res = c()
for (m in colnames(dti_tracts)[56:ncol(dti_tracts)]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(dti_tracts[dti_tracts$age_at_scan < 18, ], 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
write.csv(res, file='~/data/prs/results_PRSp3_DTI_DX_histogramCleanedWithinTract2SD_AgeSE18.csv', row.names=F)
```

```{r}
res = c()
for (m in colnames(dti_tracts)[56:ncol(dti_tracts)]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(dti_tracts[dti_tracts$age_at_scan < 16, ], 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
write.csv(res, file='~/data/prs/results_PRSp3_DTI_DX_histogramCleanedWithinTract2SD_AgeSE16.csv', row.names=F)
```

```{r}
res = c()
for (m in colnames(dti_tracts)[56:ncol(dti_tracts)]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(dti_tracts[dti_tracts$age_at_scan < 12, ], 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
write.csv(res, file='~/data/prs/results_PRSp3_DTI_DX_histogramCleanedWithinTract2SD_AgeSE12.csv', row.names=F)
```

None of these look good... let's see what SPSS does with this cleaned dataset:

```{r}
write.csv(dti_tracts, row.names=F, file='/Volumes/Shaw/prs_mediation/spss_dti_cleanedHistogram_within2SD.csv')
```

Now we'll make the mediation function a bit fancier:

```{r}
library(mediation)
run_mediation = function(data, X, M, Y, nboot=100, short=T) {
    data = data[!is.na(data[, M]), ]
    data[, c(X, M)] = scale(data[, c(X, M)])
    data = as.data.frame(data)
    # just removing the "exclude" factor
    data$DX = factor(data$ADHD_current_yes_no)
    Y = 'DX'
    
    fm = as.formula(sprintf('%s ~ %s + Sex + age', M, X))
    fy = as.formula(sprintf('%s ~ %s + %s + Sex + age', Y, X, M))
    model.M <- lm(fm, data=data)
    model.Y <- glm(fy, data=data, family=binomial(link='logit'))
    results <- mediate(model.M, model.Y, treat=X, mediator=M, boot=T, sims=nboot, boot.ci.type='bca')
    fit = lm(as.formula(sprintf('%s ~ %s + Sex + age', X, Y)), data=data)
    fit2 = lm(as.formula(sprintf('%s ~ %s + Sex + age', M, Y)), data=data)
    if (short) {
      mycoef = which(rownames(summary(model.M)$coefficients) == X)
      tstat.M <- summary(model.M)$coefficients[mycoef, 3]
      pval.M <- summary(model.M)$coefficients[mycoef, 4]
      mycoef = which(rownames(summary(fit)$coefficients) == 'DXyes')
      tstat <- summary(fit)$coefficients[mycoef, 3]
      pval <- summary(fit)$coefficients[mycoef, 4]
      mycoef = which(rownames(summary(fit2)$coefficients) == 'DXyes')
      tstat2 <- summary(fit2)$coefficients[mycoef, 3]
      pval2 <- summary(fit2)$coefficients[mycoef, 4]
      res = c(results$mediator, results$nobs, results$tau.coef, results$tau.p, results$d.avg, results$d.avg.p,
              results$z.avg, results$z.avg.p, results$n.avg, results$n.avg.p, tstat.M, pval.M, tstat, pval, tstat2, pval2)
      names(res) = c('M', 'nobs', 'tot', 'tot_p', 'acme', 'acme_p', 'ade', 'ade_p', 'prop', 'prop_p',
                     'XtoM', 'XtoM_p', 'XtoYnoM', 'XtoYnoM_p', 'MtoYnoX', 'MtoYnoX_p')
      return(res)
    } else {
      return(results)
    }
}
```

Now we need to make sure the age is the actual age_at_scan, and re-clean everything:

```{r}
rm_me = tracts$numVolumes < 60 | tracts$norm.rot > .003 | tracts$norm.trans > .3 | tracts$mean_fa < .3 | tracts$mean_ad < .97 | tracts$mean_rd < .5
dti_tracts = tracts[!rm_me, ]
dti_tracts$age = dti_tracts$age_at_scan
phen_vars = c(which(grepl("^FA_", colnames(dti_tracts))),
              which(grepl("^AD_", colnames(dti_tracts))),
              which(grepl("^RD_", colnames(dti_tracts))),
              which(grepl("^MO_", colnames(dti_tracts))))
for (j in phen_vars) {
  upper = mean(dti_tracts[, j], na.rm=T) + 2*sd(dti_tracts[, j], na.rm=T)
  lower = mean(dti_tracts[, j], na.rm=T) - 2*sd(dti_tracts[, j], na.rm=T)
  idx = which(dti_tracts[, j] < lower | dti_tracts[, j] > upper)
  dti_tracts[idx, j] = NA
}
res = c()
for (m in colnames(dti_tracts)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(dti_tracts, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
write.csv(res, file='~/data/prs/results_PRSp3_DTI_DX_PlusSexPlusAge_histogramCleanedWithinTract2SD.csv', row.names=F)
```

# Structural

Let's incorporate some of the structural QC as well:

```{r}
struct = read.csv('~/data/prs/struct_07112017.csv')
rois = merge(df, struct, by='MRN')
keep_me = which(rois$avg_freesurfer_score < 2.5 & rois$MPRAGE_QC <= 2)
struct_rois = rois[keep_me, ]

phen_vars = c(which(grepl("^lh_", colnames(struct_rois))),
              which(grepl("^rh_", colnames(struct_rois))),
              250:311)
for (j in phen_vars) {
  upper = mean(struct_rois[, j], na.rm=T) + 2*sd(struct_rois[, j], na.rm=T)
  lower = mean(struct_rois[, j], na.rm=T) - 2*sd(struct_rois[, j], na.rm=T)
  idx = which(struct_rois[, j] < lower | struct_rois[, j] > upper)
  struct_rois[idx, j] = NA
}
res = c()
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
write.csv(res, file='~/data/prs/results_PRSp3_ROIs_DX_PlusSexPlusAge_histogramCleanedWithinTract2SD.csv', row.names=F)

```

Let's make the function even more flexible, so that we can change the covariates as we go and correct for overall brain measures:

```{r}
library(mediation)
run_mediation = function(data, X, M, Y, covars, nboot=100, short=T) {
    data = data[!is.na(data[, M]), ]
    data[, c(X, M)] = scale(data[, c(X, M)])
    data = as.data.frame(data)
    # just removing the "exclude" factor
    data$DX = factor(data$ADHD_current_yes_no)
    Y = 'DX'
    
    fm = as.formula(sprintf('%s ~ %s + %s', M, X, covars))
    fy = as.formula(sprintf('%s ~ %s + %s + %s', Y, X, M, covars))
    model.M <- lm(fm, data=data)
    model.Y <- glm(fy, data=data, family=binomial(link='logit'))
    results <- mediate(model.M, model.Y, treat=X, mediator=M, boot=T, sims=nboot, boot.ci.type='bca')
    fit = lm(as.formula(sprintf('%s ~ %s + %s', X, Y, covars)), data=data)
    fit2 = lm(as.formula(sprintf('%s ~ %s + %s', M, Y, covars)), data=data)
    if (short) {
      mycoef = which(rownames(summary(model.M)$coefficients) == X)
      tstat.M <- summary(model.M)$coefficients[mycoef, 3]
      pval.M <- summary(model.M)$coefficients[mycoef, 4]
      mycoef = which(rownames(summary(fit)$coefficients) == 'DXyes')
      tstat <- summary(fit)$coefficients[mycoef, 3]
      pval <- summary(fit)$coefficients[mycoef, 4]
      mycoef = which(rownames(summary(fit2)$coefficients) == 'DXyes')
      tstat2 <- summary(fit2)$coefficients[mycoef, 3]
      pval2 <- summary(fit2)$coefficients[mycoef, 4]
      res = c(results$mediator, results$nobs, results$tau.coef, results$tau.p, results$d.avg, results$d.avg.p,
              results$z.avg, results$z.avg.p, results$n.avg, results$n.avg.p, tstat.M, pval.M, tstat, pval, tstat2, pval2)
      names(res) = c('M', 'nobs', 'tot', 'tot_p', 'acme', 'acme_p', 'ade', 'ade_p', 'prop', 'prop_p',
                     'XtoM', 'XtoM_p', 'XtoYnoM', 'XtoYnoM_p', 'MtoYnoX', 'MtoYnoX_p')
      return(res)
    } else {
      return(results)
    }
}
```

Now we have to do this by overall structure type:

```{r}
struct = read.csv('~/data/prs/struct_07112017.csv')
rois = merge(df, struct, by='MRN')
keep_me = which(rois$avg_freesurfer_score < 2.5 & rois$MPRAGE_QC <= 2)
struct_rois = rois[keep_me, ]

phen_vars = c(which(grepl("^lh_", colnames(struct_rois))),
              which(grepl("^rh_", colnames(struct_rois))),
              250:311)
for (j in phen_vars) {
  upper = mean(struct_rois[, j], na.rm=T) + 2*sd(struct_rois[, j], na.rm=T)
  lower = mean(struct_rois[, j], na.rm=T) - 2*sd(struct_rois[, j], na.rm=T)
  idx = which(struct_rois[, j] < lower | struct_rois[, j] > upper)
  struct_rois[idx, j] = NA
}
res = c()
phen_vars = which(grepl("lh_[a-z]*_thickness", colnames(struct_rois)))
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no', 'Sex + age + lh_MeanThickness_thickness'))
}
phen_vars = which(grepl("rh_[a-z]*_thickness", colnames(struct_rois)))
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no', 'Sex + age + rh_MeanThickness_thickness'))
}
phen_vars = which(grepl("_volume$", colnames(struct_rois)))
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no', 'Sex + age + EstimatedTotalIntraCranialVol'))
}
phen_vars = which(grepl("lh_[a-z]*_area", colnames(struct_rois)))
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no', 'Sex + age + lh_WhiteSurfArea_area'))
}
phen_vars = which(grepl("rh_[a-z]*_area", colnames(struct_rois)))
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no', 'Sex + age + rh_WhiteSurfArea_area'))
}
write.csv(res, file='~/data/prs/results_PRSp3_ROIs_DX_PlusSexPlusAgePlusTotal_histogramCleanedWithinTract2SD.csv', row.names=F)
```

Let's run just the kids now:

```{r}
struct = read.csv('~/data/prs/struct_07112017.csv')
rois = merge(df, struct, by='MRN')
keep_me = which(rois$avg_freesurfer_score < 2.5 & rois$MPRAGE_QC <= 2)
struct_rois = rois[keep_me, ]

phen_vars = c(which(grepl("^lh_", colnames(struct_rois))),
              which(grepl("^rh_", colnames(struct_rois))),
              250:311)
for (j in phen_vars) {
  upper = mean(struct_rois[, j], na.rm=T) + 2*sd(struct_rois[, j], na.rm=T)
  lower = mean(struct_rois[, j], na.rm=T) - 2*sd(struct_rois[, j], na.rm=T)
  idx = which(struct_rois[, j] < lower | struct_rois[, j] > upper)
  struct_rois[idx, j] = NA
}
struct_rois = struct_rois[struct_rois$age <= 18, ]
res = c()
phen_vars = which(grepl("lh_[a-z]*_thickness", colnames(struct_rois)))
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no', 'Sex + age + lh_MeanThickness_thickness'))
}
phen_vars = which(grepl("rh_[a-z]*_thickness", colnames(struct_rois)))
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no', 'Sex + age + rh_MeanThickness_thickness'))
}
phen_vars = which(grepl("_volume$", colnames(struct_rois)))
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no', 'Sex + age + EstimatedTotalIntraCranialVol'))
}
phen_vars = which(grepl("lh_[a-z]*_area", colnames(struct_rois)))
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no', 'Sex + age + lh_WhiteSurfArea_area'))
}
phen_vars = which(grepl("rh_[a-z]*_area", colnames(struct_rois)))
for (m in colnames(struct_rois)[phen_vars]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(struct_rois, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no', 'Sex + age + rh_WhiteSurfArea_area'))
}
write.csv(res, file='~/data/prs/results_PRSp3_ROIs_DX_PlusSexPlusAgePlusTotal_histogramCleanedWithinTract2SD_AgeSE18.csv', row.names=F)
```

* Other PRS scores?
* ICA for resting state, using just the default mode one
* Brain space?






