---
title: "Mediation through different variables"
output: html_notebook
---

Now that we know that ADHD_DX is related to many of the PRS thresholds, let's see how different variables mediate the result:

```{r}
gf = read.csv('~/data/prs/clinical_06192017.csv')
gf = gf[gf$ADHD_current_yes_no!='exclude',]
pgc = read.csv('~/data/prs/PRS2017_original_clump_default.csv')
```

Let's combine the clinical data and PRS to restrict the size of our matrices:

```{r}
df = merge(gf, pgc)
# remove duplicated MRNs
df = df[!duplicated(df$MRN),]
```

Now we're ready to run our LME using family ID as the random term, so we can have an idea of what threshold to use:

```{r}
library(nlme)

print(sprintf('NV vs ADHD: %d subjects', sum(!is.na(df$ADHD_current_yes_no))))
# start with quick T-test for DX
vars = which(grepl('^PROFILES', colnames(df)))
all_res = c()
x_str = 'ADHD_current_yes_no'
for (i in 1:length(vars)) {
    y_str = colnames(df)[vars][i]
    fm_str = sprintf('%s ~ %s', y_str, x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df, na.action = na.omit))
    last_row = nrow(summary(fit)$tTable)
    res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    all_res = rbind(all_res, res)
}
colnames(all_res) = c(sprintf('Tstat_%s', x_str), sprintf('pval_%s', x_str))
rownames(all_res) = colnames(df)[vars]
print(all_res)
```

Let's go with .3 for now. We'll try a simple mediation model, using IQ because we know it works:

```{r}
X = 'PROFILES.0.3.profile'
M = 'IQ'
Y = 'ADHD_current_yes_no'
library(mediation)
neuropsych = read.csv('~/data/prs/neuropsych_07072017.csv')
data = merge(df, neuropsych, by='MRN')
data = data[!is.na(data$IQ), ]

fm = as.formula(sprintf('%s ~ %s', M, X))
fy = as.formula(sprintf('%s ~ %s + %s', Y, X, M))
model.M <- lm(fm, data=data)
model.Y <- glm(fy, data=data, family=binomial(link='logit'))
results <- mediate(model.M, model.Y, treat=X, mediator=M, boot=T, sims=100, boot.ci.type='bca')
summary(results)
```

What are the effects of scaling the variables?

```{r}
X = 'PROFILES.0.3.profile'
M = 'IQ'
Y = 'ADHD_current_yes_no'
library(mediation)
neuropsych = read.csv('~/data/prs/neuropsych_07072017.csv')
data = merge(df, neuropsych, by='MRN')
data = data[!is.na(data$IQ), ]
data[, c(X, M)] = scale(data[, c(X, M)])
data = as.data.frame(data)

fm = as.formula(sprintf('%s ~ %s', M, X))
fy = as.formula(sprintf('%s ~ %s + %s', Y, X, M))
model.M <- lm(fm, data=data)
model.Y <- glm(fy, data=data, family=binomial(link='logit'))
results <- mediate(model.M, model.Y, treat=X, mediator=M, boot=T, sims=100, boot.ci.type='bca')
summary(results)
plot(results)
```

These seem to be quite different than the ones without scaling the variables. What's up?

First, let's interpret this scaled model, because the numbers seem to make a bit more sense. As we're using linear models, Total Effect ~ ADE + ACME. This *Total effect* is the effect of X in Y without M (PRS on DX without IQ). The *Direct Effect (ADE)* is the direct effect of X on Y after taking into account a mediation (indirect) effect of M (i.e. direct effect of PRS on DX after taking IQ into account). Finally, the *Average Causal Mediation Effects (ACME)* is the total minus the direct effect. The goal of mediation analysis is to obtain this indirect effect and see if it’s statistically significant, which in our case actually is.

The results above could be seen as: of the estimated -.06735 (the Total Effect) decrease in likelihood of ADHD DX due to PRS, an estimated -0.01874 (ACME (average)) is as a result of IQ and the remaining -0.04860 (ADE (average)) is from PRS itself. Since the outcome (DX) is binary all estimated effects are expressed as the increase/decrease in probability that the subject has ADHD. Also note that the difference in mediation between the two classes is negligible, so it's OK to average them.

Now, what's going on with scaling? Maybe it's because it uses the actual coefficients in the equations. So, if we're looking at a correlation, the coefficients need to be standardized to match something like R. If X were binary, it wouldn't matter. In any case, let's keep on scaling them.

Now, let's see how it compares to SPSS:

```{r}
write.csv(data, file='/Volumes/Shaw/prs_mediation/spss_test.csv')
```

```
Run MATRIX procedure:

************** PROCESS Procedure for SPSS Release 2.16.1 *****************

          Written by Andrew F. Hayes, Ph.D.       www.afhayes.com
    Documentation available in Hayes (2013). www.guilford.com/p/hayes3

**************************************************************************
Model = 4
    Y = ADHDyn
    X = PRSp3
    M = IQ

Sample size
        953

**************************************************************************
Outcome: IQ

Model Summary
          R       R-sq        MSE          F        df1        df2          p
      .2733      .0747      .9263    76.7582     1.0000   951.0000      .0000

Model
              coeff         se          t          p       LLCI       ULCI
constant      .0000      .0312      .0000     1.0000     -.0612      .0612
PRSp3         .2733      .0312     8.7612      .0000      .2121      .3345

**************************************************************************
**************************************************************************
Outcome: ADHDyn

Coding of binary DV for analysis:
    ADHDyn  Analysis
       .00       .00
       .00      1.00

Logistic Regression Summary
       -2LL   Model LL    p-value   McFadden   CoxSnell   Nagelkrk          n
  1217.5486    38.8354      .0000      .0309      .0399      .0545   953.0000

Model
              coeff         se          Z          p       LLCI       ULCI
constant     -.5506      .0687    -8.0088      .0000     -.6854     -.4159
IQ           -.3127      .0725    -4.3129      .0000     -.4548     -.1706
PRSp3        -.2175      .0709    -3.0664      .0022     -.3565     -.0785

******************** DIRECT AND INDIRECT EFFECTS *************************

Direct effect of X on Y
     Effect         SE          Z          p       LLCI       ULCI
     -.2175      .0709    -3.0664      .0022     -.3565     -.0785

Indirect effect of X on Y
       Effect    Boot SE   BootLLCI   BootULCI
IQ     -.0855      .0221     -.1335     -.0466
******************** ANALYSIS NOTES AND WARNINGS *************************

Number of bootstrap samples for bias corrected bootstrap confidence intervals:
     5000

Level of confidence for all confidence intervals in output:
    95.00

NOTE: The TOTAL option is not available when Y is dichotomous.

------ END MATRIX -----
```

The numbers don't match perfectly, btu at least in overall significance we're in the same ballpark. Let me try to increase the number of bootstraps:

```{r}
X = 'PROFILES.0.3.profile'
M = 'IQ'
Y = 'ADHD_current_yes_no'
library(mediation)
neuropsych = read.csv('~/data/prs/neuropsych_07072017.csv')
data = merge(df, neuropsych, by='MRN')
data = data[!is.na(data$IQ), ]
data[, c(X, M)] = scale(data[, c(X, M)])
data = as.data.frame(data)

fm = as.formula(sprintf('%s ~ %s', M, X))
fy = as.formula(sprintf('%s ~ %s + %s', Y, X, M))
model.M <- lm(fm, data=data)
model.Y <- glm(fy, data=data, family=binomial(link='logit'))
results <- mediate(model.M, model.Y, treat=X, mediator=M, boot=T, sims=5000, boot.ci.type='bca')
summary(results)
```

As expected, the results just get stronger. For better comparison with SPSS, here's the logit model:

```{r}
summary(model.Y)
```

So, this matches quite well. The main differences then are in the calculation of the mediation. But it doesn't look like the differences are that big anyways.

But the mediation package can also run lmer, even though bootstrap doesn't work... should we try it?

```{r}
library(lme4)

X = 'PROFILES.0.3.profile'
M = 'IQ'
Y = 'ADHD_current_yes_no'
library(mediation)
neuropsych = read.csv('~/data/prs/neuropsych_07072017.csv')
data = merge(df, neuropsych, by='MRN')
data = data[!is.na(data$IQ), ]
# here we have to scale otherwise we get some weird warnings
data[, c(X, M)] = scale(data[, c(X, M)])

fm = as.formula(sprintf('%s ~ %s + (1|NuclearFamID)', M, X))
fy = as.formula(sprintf('%s ~ %s + %s + (1|NuclearFamID)', Y, X, M))
model.M <- lmer(fm, data=data)
model.Y <- glmer(fy, data=data, family=binomial(link='probit'))
results <- mediate(model.M, model.Y, treat=X, mediator=M, sims=100)
summary(results)
plot(results)
```

The results using mixed-effect models is similar to not using it, which is reassuring. Also, if we use simpler models (i.e. non-mixed), we could also run some functions from the MBESS package, which would allow us to calculate effect sizes for the estimates.

# Checking several different mediators

Now we can have some fun checking several different mediators.

```{r}
library(mediation)
run_mediation = function(data, X, M, Y, nboot=100, short=T) {
    data = data[!is.na(data[, M]), ]
    data[, c(X, M)] = scale(data[, c(X, M)])
    data = as.data.frame(data)
    
    fm = as.formula(sprintf('%s ~ %s', M, X))
    fy = as.formula(sprintf('%s ~ %s + %s', Y, X, M))
    model.M <- lm(fm, data=data)
    model.Y <- glm(fy, data=data, family=binomial(link='logit'))
    results <- mediate(model.M, model.Y, treat=X, mediator=M, boot=T, sims=nboot, boot.ci.type='bca')
    if (short) {
      res = c(results$mediator, results$nobs, results$tau.coef, results$tau.p, results$d.avg, results$d.avg.p,
              results$z.avg, results$z.avg.p, results$n.avg, results$n.avg.p)
      names(res) = c('M', 'nobs', 'tot', 'tot_p', 'acme', 'acme_p', 'ade', 'ade_p', 'prop', 'prop_p')
      return(res)
    } else {
      return(results)
    }
}
```

Now we're ready to try different variables:

```{r}
neuropsych = read.csv('~/data/prs/neuropsych_07072017.csv')
data = merge(df, neuropsych, by='MRN')
res = c()
for (m in colnames(neuropsych)[2:ncol(neuropsych)]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(data, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
res = cbind(res, 'neuropsych')
colnames(res)[ncol(res)] = 'dtype'
all_res = res
```
```{r}
dti = read.csv('~/data/prs/dti_07062017.csv')
data = merge(df, dti, by='MRN')
res = c()
for (m in colnames(dti)[25:ncol(dti)]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(data, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
res = cbind(res, 'dti')
colnames(res)[ncol(res)] = 'dtype'
all_res = rbind(all_res, res) 
```

```{r}
struct = read.csv('~/data/prs/struct_07072017_v3.csv')
data = merge(df, struct, by='MRN')
res = c()
for (m in colnames(struct)[11:ncol(struct)]) {
  print(sprintf('Running %s', m))
  res = rbind(res, run_mediation(data, 'PROFILES.0.3.profile', m, 'ADHD_current_yes_no'))
}
res = cbind(res, 'struct')
colnames(res)[ncol(res)] = 'dtype'
all_res = rbind(all_res, res) 
```







