---
title: "Clinical to PRS"
output: html_notebook
---

Here we are trying to use PRS to predict any of the clinical variables. If we get a good correlation (or T-test), then we can try some sort of mediation analysis using other variables as the mediator (e.g. brain, cognition) and making the correlation go away.

Note that this clinical file we're loading will keep on getting updated, cleaned, etc. So, let's make sure to save different versions of this analysis.

```{r}
gf = read.csv('~/data/prs/clinical_06192017.csv')
gf = gf[gf$ADHD_current_yes_no!='exclude',]
pgc = read.csv('~/data/prs/PRS2017_noInversion_all.csv')
# pgc = read.csv('~/data/prs/PRS2017_noInversion_all.csv')
pgc_eur = read.csv('~/data/prs/PRS2017_noInversion_eur_all.csv')
```

Let's combine the clinical data and PRS to restrict the size of our matrices:

```{r}
df = merge(gf, pgc)
# remove duplicated MRNs
df = df[!duplicated(df$MRN),]
df_eur = merge(gf, pgc_eur)
df_eur = df_eur[!duplicated(df_eur$MRN),]
eur_idx = df_eur$Race=='[White]' & df_eur$Ethnicity=='Not Hispanic or Latino'
df_eur = df_eur[eur_idx, ]
```

Now we're ready to run our LME using family ID as the random term:

```{r}
library(nlme)

print(sprintf('NV vs ADHD: %d subjects', sum(!is.na(df$ADHD_current_yes_no))))
# start with quick T-test for DX
vars = which(grepl('^PROFILES', colnames(df)))
all_res = c()
x_str = 'ADHD_current_yes_no'
for (i in 1:length(vars)) {
    y_str = colnames(df)[vars][i]
    fm_str = sprintf('%s ~ %s', y_str, x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df, na.action = na.omit))
    last_row = nrow(summary(fit)$tTable)
    res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    all_res = rbind(all_res, res)
}
colnames(all_res) = c(sprintf('Tstat_%s', x_str), sprintf('pval_%s', x_str))
rownames(all_res) = colnames(df)[vars]
print(all_res)
```

We also do some current SX regression only within ADHDs:

```{r}
adhd_idx = df$ADHD_current_yes_no=='yes'
print(sprintf('ADHD-only regressions: %d subjects', sum(adhd_idx)))
targets = c('SX_inatt', 'SX_HI')
for (y_str in targets) {
  sx_res = c()
  for (i in 1:length(vars)) {
    x_str = colnames(df)[vars][i]
    fm_str = sprintf('%s ~ %s', y_str, x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df[adhd_idx,], na.action = na.omit))
    last_row = nrow(summary(fit)$tTable)
    res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    sx_res = rbind(sx_res, res)
  }
  colnames(sx_res) = c(sprintf('Tstat_%s_adhdOnly', y_str), sprintf('pval_%s_adhdOnly', y_str))
  rownames(sx_res) = colnames(df)[vars]
  all_res = cbind(all_res, sx_res)
  print(sx_res)
}
```

And regression for worst SX symptoms (lifetime):

```{r}
adhd_idx = df$ADHD_current_yes_no=='yes'
targets = c('inatt', 'HI')
for (y_str in targets) {
  y = pmax(df[adhd_idx, sprintf('SX_%s', y_str)],
           df[adhd_idx, sprintf('%s_child', y_str)],
           na.rm=T)
  sx_res = c()
  for (i in 1:length(vars)) {
    x_str = colnames(df)[vars][i]
    fm_str = sprintf('y ~ %s', x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df[adhd_idx,], na.action = na.omit))
    last_row = nrow(summary(fit)$tTable)
    res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    sx_res = rbind(sx_res, res)
  }
  colnames(sx_res) = c(sprintf('Tstat_Max%s_adhdOnly', y_str), sprintf('pval_Max%s_adhdOnly', y_str))
  rownames(sx_res) = colnames(df)[vars]
  all_res = cbind(all_res, sx_res)
  print(sx_res)
}
```

Let's do both regressions above but including NVs. Here, we assume any NVs without SX to be 0:

```{r}
assume_idx = df$ADHD_current_yes_no=='no' & is.na(df$SX_HI + df$SX_inatt)
df2 = df
df2[assume_idx, 'SX_inatt'] = 0
df2[assume_idx, 'SX_HI'] = 0
targets = c('SX_inatt', 'SX_HI')
for (y_str in targets) {
  sx_res = c()
  for (i in 1:length(vars)) {
    x_str = colnames(df2)[vars][i]
    fm_str = sprintf('%s ~ %s', y_str, x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df2, na.action = na.omit))
    if (length(fit) > 1) {
      last_row = nrow(summary(fit)$tTable)
      res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    } else {
      res = c(NA, NA)
    }
    sx_res = rbind(sx_res, res)
  }
  colnames(sx_res) = c(sprintf('Tstat_%s', y_str), sprintf('pval_%s', y_str))
  rownames(sx_res) = colnames(df)[vars]
  all_res = cbind(all_res, sx_res)
  print(sx_res)
}

targets = c('inatt', 'HI')
for (y_str in targets) {
  y = pmax(df2[, sprintf('SX_%s', y_str)],
           df2[, sprintf('%s_child', y_str)],
           na.rm=T)
  sx_res = c()
  for (i in 1:length(vars)) {
    x_str = colnames(df2)[vars][i]
    fm_str = sprintf('y ~ %s', x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df2, na.action = na.omit))
    last_row = nrow(summary(fit)$tTable)
    res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    sx_res = rbind(sx_res, res)
  }
  colnames(sx_res) = c(sprintf('Tstat_Max%s', y_str), sprintf('pval_Max%s', y_str))
  rownames(sx_res) = colnames(df)[vars]
  all_res = cbind(all_res, sx_res)
  print(sx_res)
}
```

Let's write out the table for easier visualization:

```{r}
write.csv(all_res, file='~/philip/prs_results_noInversion_09202017.csv')
```

# European-only sample
And now we do the exact same analysis, but using the European-only sample

```{r}
print(sprintf('NV vs ADHD: %d subjects', sum(!is.na(df_eur$ADHD_current_yes_no))))
# start with quick T-test for DX
vars = which(grepl('^PROFILES', colnames(df_eur)))
all_res = c()
x_str = 'ADHD_current_yes_no'
for (i in 1:length(vars)) {
    y_str = colnames(df_eur)[vars][i]
    fm_str = sprintf('%s ~ %s', y_str, x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df_eur, na.action = na.omit))
    last_row = nrow(summary(fit)$tTable)
    res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    all_res = rbind(all_res, res)
}
colnames(all_res) = c(sprintf('Tstat_%s', x_str), sprintf('pval_%s', x_str))
rownames(all_res) = colnames(df_eur)[vars]

adhd_idx = df_eur$ADHD_current_yes_no=='yes'
print(sprintf('ADHD-only regressions: %d subjects', sum(adhd_idx)))
targets = c('SX_inatt', 'SX_HI')
for (y_str in targets) {
  sx_res = c()
  for (i in 1:length(vars)) {
    x_str = colnames(df_eur)[vars][i]
    fm_str = sprintf('%s ~ %s', y_str, x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df_eur[adhd_idx,], na.action = na.omit))
    last_row = nrow(summary(fit)$tTable)
    res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    sx_res = rbind(sx_res, res)
  }
  colnames(sx_res) = c(sprintf('Tstat_%s_adhdOnly', y_str), sprintf('pval_%s_adhdOnly', y_str))
  rownames(sx_res) = colnames(df_eur)[vars]
  all_res = cbind(all_res, sx_res)
}

adhd_idx = df$ADHD_current_yes_no=='yes'
targets = c('inatt', 'HI')
for (y_str in targets) {
  y = pmax(df[adhd_idx, sprintf('SX_%s', y_str)],
           df[adhd_idx, sprintf('%s_child', y_str)],
           na.rm=T)
  sx_res = c()
  for (i in 1:length(vars)) {
    x_str = colnames(df_eur)[vars][i]
    fm_str = sprintf('y ~ %s', x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df_eur[adhd_idx,], na.action = na.omit))
    last_row = nrow(summary(fit)$tTable)
    res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    sx_res = rbind(sx_res, res)
  }
  colnames(sx_res) = c(sprintf('Tstat_Max%s_adhdOnly', y_str), sprintf('pval_Max%s_adhdOnly', y_str))
  rownames(sx_res) = colnames(df_eur)[vars]
  all_res = cbind(all_res, sx_res)
}

assume_idx = df_eur$ADHD_current_yes_no=='no' & is.na(df_eur$SX_HI + df_eur$SX_inatt)
df2 = df_eur
df2[assume_idx, 'SX_inatt'] = 0
df2[assume_idx, 'SX_HI'] = 0
targets = c('SX_inatt', 'SX_HI')
for (y_str in targets) {
  sx_res = c()
  for (i in 1:length(vars)) {
    x_str = colnames(df2)[vars][i]
    fm_str = sprintf('%s ~ %s', y_str, x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df2, na.action = na.omit))
    if (length(fit) > 1) {
      last_row = nrow(summary(fit)$tTable)
      res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    } else {
      res = c(NA, NA)
    }
    sx_res = rbind(sx_res, res)
  }
  colnames(sx_res) = c(sprintf('Tstat_%s', y_str), sprintf('pval_%s', y_str))
  rownames(sx_res) = colnames(df_eur)[vars]
  all_res = cbind(all_res, sx_res)
}

targets = c('inatt', 'HI')
for (y_str in targets) {
  y = pmax(df2[, sprintf('SX_%s', y_str)],
           df2[, sprintf('%s_child', y_str)],
           na.rm=T)
  sx_res = c()
  for (i in 1:length(vars)) {
    x_str = colnames(df2)[vars][i]
    fm_str = sprintf('y ~ %s', x_str) # make sure the interesting variable comes last!
    fit = try(lme(as.formula(fm_str), random=~1|NuclearFamID, data=df2, na.action = na.omit))
    last_row = nrow(summary(fit)$tTable)
    res = c(summary(fit)$tTable[last_row,'t-value'], summary(fit)$tTable[last_row,'p-value'])
    sx_res = rbind(sx_res, res)
  }
  colnames(sx_res) = c(sprintf('Tstat_Max%s', y_str), sprintf('pval_Max%s', y_str))
  rownames(sx_res) = colnames(df_eur)[vars]
  all_res = cbind(all_res, sx_res)
}

write.csv(all_res, file='~/philip/prs_eur_results_noInversion_09202017.csv')

```

# TODO:
* clean up clinical file
* play with age groups
* run mediation analysis for good correlations
